---
title: "Seasonality_11_June_2026"
output: html_document
date: "2026-June-11"
---

### Import libraries
```{r}
library(haven)
library(ggplot2)
library(ggfortify)
library(lubridate)
library(dplyr)
library(ggbeeswarm)
library(lme4) 
library(lmerTest)
library(car)
library(ggpubr)
library(emmeans)
library(sjstats)
library(mgcv)
library(tidyr)
library(visreg)
library(itsadug)
library(gratia)
library(readr)
library(glue)
library(purrr)
library(scales)
library(stringr)
library(metR)
library(broom)
library(hms)
library(ggpattern)
library(patchwork)
library(ggpp)
library(ggsignif)

# ---- Cyclic spline knots: enforce a true 12-month period ----
# With bs="cc" and MONTH_NUM in 1:12, the default knot placement ties
# month 1 == month 12, giving an 11-month cycle that forces Dec == Jan and
# distorts the December peak and every seasonal amplitude. Boundary knots at
# 0.5 and 12.5 give a proper 12-month period (Dec -> Jan is one step).
# Passed via `knots = MONTH_KNOTS` to every bam()/gam() with a MONTH_NUM cyclic smooth.
MONTH_KNOTS <- list(MONTH_NUM = c(0.5, 12.5))
```

### Read in data
```{r}
# Use cached RDS if available; otherwise read CSV and cache
data_rds <- "/Users/gregory.grosicki/Desktop/Seasonality/Seasonality_npjDig/Seasonality_Pull_22_Apr_2026.rds"
data_csv <- "/Users/gregory.grosicki/Desktop/Seasonality/Seasonality_npjDig/Seasonality_Pull_22_Apr_2026.csv"

if (file.exists(data_rds)) {
  df <- readRDS(data_rds)
} else {
  df <- read.csv(data_csv)
  saveRDS(df, data_rds)
}

# Create field for BMI
df <- df %>%
  mutate(BMI = WEIGHT / (HEIGHT^2))

# Create field for SHRZS
df <- df %>%
  mutate(
    SHRZS = TOTAL_ZONE1_MIN * 1 +
            TOTAL_ZONE2_MIN * 2 +
            TOTAL_ZONE3_MIN * 3 +
            TOTAL_ZONE4_MIN * 4 +
            TOTAL_ZONE5_MIN * 5
  )
df <- df %>%
  mutate(SHRZS = ifelse(is.na(SHRZS), 0, SHRZS))
```

### Setup: Color palettes, theme, and helper functions
```{r}
# ---- Color palettes ----
age_colors <- c("Young" = "#66c2a5", "Middle" = "#fc8d62", "Old" = "#8da0cb")
gender_colors <- c("male" = "#1f78b4", "female" = "#f781bf")
gender_labels <- c("male" = "Male", "female" = "Female")
pa_colors <- c("Low" = "#e41a1c", "Moderate" = "#377eb8", "High" = "#4daf4a")
bmi_colors <- c("Normal" = "#4daf4a", "Overweight" = "#377eb8", "Obese" = "#e41a1c")

# ---- Custom theme ----
theme_seasonality <- function(legend = TRUE) {
  t <- theme_minimal(base_size = 14) +
    theme(
      axis.line = element_line(color = "black", linewidth = 0.8),
      axis.text = element_text(color = "black", face = "bold"),
      axis.title = element_text(color = "black", face = "bold"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      plot.title = element_blank(),
      panel.border = element_blank()
    )
  if (!legend) {
    t <- t + theme(legend.position = "none")
  } else {
    t <- t + theme(
      legend.title = element_text(color = "black", face = "bold"),
      legend.text = element_text(color = "black", face = "bold")
    )
  }
  t
}

# ---- Bootstrap: participant-level resampling ----

# Single or grouped amplitude bootstrap
bootstrap_amplitude <- function(data, formula, by = NULL, B = 1000, seed = 2024) {
  set.seed(seed)
  user_ids <- unique(data$USER_ID)
  n_users <- length(user_ids)
  results <- vector("list", B)

  for (i in 1:B) {
    boot_ids <- sample(user_ids, n_users, replace = TRUE)
    boot_data <- tibble(.boot_user = seq_along(boot_ids), USER_ID = boot_ids) %>%
      left_join(data, by = "USER_ID", relationship = "many-to-many")

    boot_model <- tryCatch(
      suppressMessages(bam(formula, data = boot_data, method = "fREML", knots = MONTH_KNOTS, discrete = TRUE)),
      error = function(e) NULL
    )

    if (is.null(boot_model)) next

    if (!is.null(by)) {
      v <- tryCatch(
        visreg(boot_model, xvar = "MONTH_NUM", by = by, overlay = TRUE,
               partial = FALSE, rug = FALSE, plot = FALSE),
        error = function(e) NULL
      )
      if (!is.null(v)) {
        results[[i]] <- v$fit %>%
          group_by(across(all_of(by))) %>%
          summarise(amplitude = max(visregFit, na.rm = TRUE) - min(visregFit, na.rm = TRUE),
                    .groups = "drop")
      }
    } else {
      v <- tryCatch(
        visreg(boot_model, xvar = "MONTH_NUM", partial = FALSE, rug = FALSE, plot = FALSE),
        error = function(e) NULL
      )
      if (!is.null(v)) {
        results[[i]] <- tibble(amplitude = max(v$fit$visregFit, na.rm = TRUE) -
                                           min(v$fit$visregFit, na.rm = TRUE))
      }
    }

    if (i %% 50 == 0) cat("Completed", i, "bootstraps\n")
  }

  bind_rows(results, .id = "iteration")
}

# Paired attenuation bootstrap (baseline vs adjusted)
bootstrap_attenuation <- function(data, formula_base, formula_adj, B = 1000, seed = 2024) {
  set.seed(seed)
  user_ids <- unique(data$USER_ID)
  n_users <- length(user_ids)
  amp_base <- rep(NA_real_, B)
  amp_adj  <- rep(NA_real_, B)

  for (i in 1:B) {
    boot_ids <- sample(user_ids, n_users, replace = TRUE)
    boot_data <- tibble(.boot_user = seq_along(boot_ids), USER_ID = boot_ids) %>%
      left_join(data, by = "USER_ID", relationship = "many-to-many")

    model_b <- tryCatch(
      bam(formula_base, data = boot_data, method = "fREML", knots = MONTH_KNOTS, discrete = TRUE),
      error = function(e) NULL
    )
    model_a <- tryCatch(
      bam(formula_adj, data = boot_data, method = "fREML", knots = MONTH_KNOTS, discrete = TRUE),
      error = function(e) NULL
    )

    if (!is.null(model_b) && !is.null(model_a)) {
      vis_b <- tryCatch(visreg(model_b, xvar = "MONTH_NUM", partial = FALSE, rug = FALSE, plot = FALSE), error = function(e) NULL)
      vis_a <- tryCatch(visreg(model_a, xvar = "MONTH_NUM", partial = FALSE, rug = FALSE, plot = FALSE), error = function(e) NULL)

      if (!is.null(vis_b) && !is.null(vis_a)) {
        amp_base[i] <- max(vis_b$fit$visregFit, na.rm = TRUE) - min(vis_b$fit$visregFit, na.rm = TRUE)
        amp_adj[i]  <- max(vis_a$fit$visregFit, na.rm = TRUE) - min(vis_a$fit$visregFit, na.rm = TRUE)
      }
    }

    if (i %% 50 == 0) cat("Completed", i, "bootstraps\n")
  }

  valid <- !is.na(amp_base) & !is.na(amp_adj)

  list(amp_base = amp_base[valid], amp_adj = amp_adj[valid], n_valid = sum(valid))
}

# Summary helper for attenuation results
summarize_attenuation <- function(att) {
  mean_base <- mean(att$amp_base)
  mean_adj  <- mean(att$amp_adj)
  mean_diff <- mean_base - mean_adj
  pct_att   <- (mean_diff / mean_base) * 100
  amp_diff  <- att$amp_base - att$amp_adj

  ci_base <- quantile(att$amp_base, c(0.025, 0.975))
  ci_adj  <- quantile(att$amp_adj, c(0.025, 0.975))
  ci_diff <- quantile(amp_diff, c(0.025, 0.975))

  # Bootstrap CI for percent attenuation (paired-resample distribution)
  pct_vec <- (att$amp_base - att$amp_adj) / att$amp_base * 100
  ci_pct  <- quantile(pct_vec, c(0.025, 0.975))

  t_result <- t.test(att$amp_base, att$amp_adj, paired = TRUE)

  cat("Mean Amplitude (Baseline):", round(mean_base, 3), "\n")
  cat("95% CI (Baseline):", round(ci_base[1], 3), "-", round(ci_base[2], 3), "\n\n")
  cat("Mean Amplitude (Adjusted):", round(mean_adj, 3), "\n")
  cat("95% CI (Adjusted):", round(ci_adj[1], 3), "-", round(ci_adj[2], 3), "\n\n")
  cat("Mean Difference:", round(mean_diff, 3), "\n")
  cat("95% CI for Difference:", round(ci_diff[1], 3), "-", round(ci_diff[2], 3), "\n\n")
  cat("Percent Attenuation:", round(pct_att, 1), "%\n")
  cat("95% CI for Percent Attenuation:", round(ci_pct[1], 1), "-", round(ci_pct[2], 1), "%\n\n")
  cat("Paired t-test p-value:", signif(t_result$p.value, 5), "\n")

  invisible(list(
    mean_base = mean_base, mean_adj = mean_adj,
    ci_base = ci_base, ci_adj = ci_adj,
    mean_diff = mean_diff, ci_diff = ci_diff,
    pct_attenuation = pct_att, ci_pct = ci_pct,
    t_test = t_result
  ))
}

# Marginal-trajectory prediction helper (lpmatrix-based)
# Computes the marginal seasonal curve averaged over factor covariates (e.g.,
# GENDER) using the proper quadratic form x'Vx for SEs. Averaging predicted
# variances across correlated predictions (as the prior version did) ignores
# Cov(p_i, p_j) and underestimates the marginal SE because predictions sharing
# fitted coefficients are highly correlated.
# Uses vcov(model, unconditional = TRUE) so CIs incorporate smoothing-parameter
# uncertainty (Bayesian credible interval; Marra & Wood 2012, Scand J Stat).
# Returns columns (MONTH_NUM, <by_var>, visregFit, visregLwr, visregUpr) — same
# shape as visreg's $fit, so downstream ggplot code is unchanged.
predict_marginal_trajectory <- function(model, data, by_var = NULL,
                                        exclude_re = NULL,
                                        month_grid = seq(1, 12, length.out = 200)) {
  pred_vars <- all.vars(delete.response(terms(model)))
  factor_marginalize <- intersect(pred_vars, c("GENDER")) |> setdiff(by_var)
  continuous <- setdiff(pred_vars, c("MONTH_NUM", by_var, "USER_ID", factor_marginalize))

  grid_args <- list(MONTH_NUM = month_grid)
  if (!is.null(by_var)) grid_args[[by_var]] <- levels(data[[by_var]])
  for (f in factor_marginalize) grid_args[[f]] <- levels(data[[f]])
  grid <- do.call(expand.grid, grid_args) |> as_tibble()

  for (v in continuous) grid[[v]] <- mean(data[[v]], na.rm = TRUE)
  if ("USER_ID" %in% pred_vars) grid$USER_ID <- data$USER_ID[1]

  # Design matrix on link scale; `exclude` zeros out columns of excluded smooths
  # (e.g. the s(USER_ID) random effect) so predictions are population-level.
  pred_args <- list(model, newdata = grid, type = "lpmatrix")
  if (!is.null(exclude_re)) pred_args$exclude <- exclude_re
  X <- do.call(predict, pred_args)

  V    <- vcov(model, unconditional = TRUE)
  beta <- coef(model)

  # Average X rows within each (MONTH_NUM, by_var) cell. The marginal
  # prediction is then x_bar' beta and its SE is sqrt(x_bar' V x_bar),
  # which correctly accounts for shared coefficient uncertainty across
  # the marginalised factor levels.
  key_cols <- c("MONTH_NUM", by_var)
  cell_id <- do.call(paste, c(as.list(grid[key_cols]), sep = ""))
  unique_cells <- unique(cell_id)
  keys_out <- grid[match(unique_cells, cell_id), key_cols, drop = FALSE] |> as_tibble()

  X_bar <- t(vapply(unique_cells,
                    function(k) colMeans(X[cell_id == k, , drop = FALSE]),
                    numeric(ncol(X))))

  fit <- as.numeric(X_bar %*% beta)
  se  <- sqrt(rowSums((X_bar %*% V) * X_bar))

  keys_out %>%
    mutate(
      visregFit = fit,
      se_avg    = se,
      visregLwr = visregFit - 1.96 * se,
      visregUpr = visregFit + 1.96 * se
    )
}

# Marginal-mean + pairwise-contrast helper (lpmatrix-based)
# Replaces the manual emm_manual + compute_all_contrasts pattern in the GAMM
# Aim 1 sections. Marginal means are computed at each (by_var, MONTH_NUM)
# cell averaging over GENDER (when GENDER is not the by_var); pairwise contrasts
# use c = X_bar_i - X_bar_j with SE = sqrt(c' V c) so the cell covariance is
# properly accounted for. Holm-adjusted across all contrasts. Uses
# vcov(model, unconditional = TRUE) for Bayesian credible intervals.
# Note: contrast direction follows factor level order in `data[[by_var]]`.
# For AGE_BUCKET this gives Young-Middle, Young-Old, Middle-Old (Y/M/O level
# order set explicitly upstream). For GENDER it follows alphabetical order
# (female - male), so signed estimates differ from prior manual versions but
# magnitudes, p-values, and CIs are identical.
emmeans_contrasts_lpmatrix <- function(model, data, by_var, months = c(8, 12),
                                       exclude_re = "s(USER_ID)") {
  pred_vars <- all.vars(delete.response(terms(model)))
  factor_marginalize <- intersect(pred_vars, c("GENDER")) |> setdiff(by_var)
  continuous <- setdiff(pred_vars, c("MONTH_NUM", by_var, "USER_ID", factor_marginalize))

  grid_args <- list(MONTH_NUM = months)
  grid_args[[by_var]] <- levels(data[[by_var]])
  for (f in factor_marginalize) grid_args[[f]] <- levels(data[[f]])
  grid <- do.call(expand.grid, grid_args) |> as_tibble()

  for (v in continuous) grid[[v]] <- mean(data[[v]], na.rm = TRUE)
  if ("USER_ID" %in% pred_vars) grid$USER_ID <- data$USER_ID[1]

  X    <- predict(model, newdata = grid, type = "lpmatrix", exclude = exclude_re)
  V    <- vcov(model, unconditional = TRUE)
  beta <- coef(model)

  cell_id <- do.call(paste, c(as.list(grid[c(by_var, "MONTH_NUM")]), sep = ""))
  unique_cells <- unique(cell_id)
  cell_meta <- grid[match(unique_cells, cell_id),
                    c(by_var, "MONTH_NUM"), drop = FALSE] |> as_tibble()
  X_bar <- t(vapply(unique_cells,
                    function(k) colMeans(X[cell_id == k, , drop = FALSE]),
                    numeric(ncol(X))))

  emm <- bind_cols(cell_meta, tibble(emmean = as.numeric(X_bar %*% beta)))

  groups <- levels(data[[by_var]])
  pairs  <- combn(groups, 2, simplify = FALSE)
  df_resid <- df.residual(model)

  contrast_rows <- list()
  for (m in months) {
    for (pair in pairs) {
      i <- which(cell_meta[[by_var]] == pair[1] & cell_meta$MONTH_NUM == m)
      j <- which(cell_meta[[by_var]] == pair[2] & cell_meta$MONTH_NUM == m)
      c_vec <- X_bar[i, ] - X_bar[j, ]
      contrast_rows[[length(contrast_rows) + 1]] <- tibble(
        MONTH_NUM = m,
        contrast  = paste(pair[1], "-", pair[2]),
        estimate  = as.numeric(c_vec %*% beta),
        SE        = sqrt(as.numeric(crossprod(c_vec, V %*% c_vec))),
        df        = df_resid
      )
    }
  }

  contrasts_df <- bind_rows(contrast_rows) %>%
    mutate(
      t.ratio  = estimate / SE,
      p.unadj  = 2 * pt(-abs(t.ratio), df = df),
      p.value  = p.adjust(p.unadj, method = "holm"),
      lower.CL = estimate - qt(0.975, df = df) * SE,
      upper.CL = estimate + qt(0.975, df = df) * SE
    ) %>%
    mutate(across(c(estimate, SE, t.ratio, p.unadj, p.value, lower.CL, upper.CL),
                  ~ round(.x, 3)))

  list(emm = emm, contrasts = contrasts_df)
}

# Pairwise bootstrap comparison helper
bootstrap_pairwise <- function(boot_df, group_var) {
  amp_wide <- boot_df %>%
    pivot_wider(names_from = all_of(group_var), values_from = amplitude) %>%
    na.omit()

  groups <- unique(boot_df[[group_var]])
  B <- nrow(amp_wide)

  pairs <- combn(groups, 2, simplify = FALSE)

  map_df(pairs, function(pair) {
    diff_vec <- amp_wide[[pair[2]]] - amp_wide[[pair[1]]]
    flips <- min(sum(diff_vec > 0), sum(diff_vec < 0))
    p_val <- if (flips == 0) {
      paste0("< ", formatC(2 / B, format = "f", digits = 3))
    } else {
      as.character(round(2 * flips / B, 3))
    }
    tibble(comparison = paste(pair[2], "vs", pair[1]), p_value = p_val)
  })
}

# ---- Inset-annotation helpers (Fig 1, 2, 3 D-I, 4, S2/S3 A&C, S6-S9) ----

# Format bootstrap p-value: pass through "< ..." strings as-is, prefix "= "
# to numeric strings. Input is the character p_value column from bootstrap_pairwise.
fmt_boot_p <- function(p) {
  ifelse(grepl("<", p), p, paste0("= ", p))
}

# Format numeric p-value (e.g. paired-t p from summarize_attenuation): "< 0.001"
# below threshold, "= 0.0XX" otherwise.
fmt_num_p <- function(p) {
  ifelse(p < 0.001, "< 0.001", sprintf("= %.3f", p))
}

# Build multi-line inset string for amplitude figures: group rows with bootstrap
# amplitude (95% CI) followed by pairwise p-values from bootstrap_pairwise.
# Headers in UPPERCASE for visual distinction (richtext/bold avoided to skip
# the gridtext dependency chain).
make_amp_inset <- function(amp_df, pval_df, by_var, group_levels = NULL) {
  if (!is.null(group_levels)) {
    amp_df <- amp_df[order(match(amp_df[[by_var]], group_levels)), ]
  }
  amp_lines <- sprintf("%s: %.2f (%.2f, %.2f)",
                       amp_df[[by_var]], amp_df$mean_amp,
                       amp_df$ci_low, amp_df$ci_high)
  pval_lines <- sprintf("%s: p %s", pval_df$comparison, fmt_boot_p(pval_df$p_value))
  paste(c("AMPLITUDE (95% CI)", amp_lines, "",
          "PAIRWISE P-VALUES", pval_lines),
        collapse = "\n")
}

# Build one-line inset string for attenuation figures (Fig 3 D-I).
make_att_inset <- function(att_summary) {
  sprintf("Attenuation: %.1f%% (%.1f, %.1f); p %s",
          att_summary$pct_attenuation,
          att_summary$ci_pct[1], att_summary$ci_pct[2],
          fmt_num_p(att_summary$t_test$p.value))
}

# Layer wrappers: drop into ggplot pipelines via `p + amp_inset_layer(...)`.
# Insets render in the right margin under the legend (npcx > 1, clip = "off").
# Each helper returns a list containing the geom + a coord_cartesian(clip = "off")
# + theme adjustments (legend anchored to top of right margin, plot margin bumped
# to make room). The helper assumes the underlying plot has no prior coord_*().
amp_inset_layer <- function(amp_df, pval_df, by_var, group_levels = NULL,
                             npcx = 1.02, npcy = 0.45, size = 2.7,
                             margin_r = 110) {
  list(
    ggpp::geom_label_npc(
      data = tibble(npcx = npcx, npcy = npcy,
                    label = make_amp_inset(amp_df, pval_df, by_var, group_levels)),
      aes(npcx = npcx, npcy = npcy, label = label),
      hjust = 0, vjust = 1, size = size, color = "black",
      label.size = 0.3, fill = scales::alpha("white", 0.85),
      label.padding = unit(0.3, "lines"), inherit.aes = FALSE
    ),
    coord_cartesian(clip = "off"),
    theme(plot.margin = margin(t = 5, r = margin_r, b = 5, l = 5),
          legend.justification = c("left", "top"))
  )
}

att_inset_layer <- function(att_summary, npcx = 1.02, npcy = 0.35, size = 3,
                             margin_r = 110) {
  list(
    ggpp::geom_label_npc(
      data = tibble(npcx = npcx, npcy = npcy, label = make_att_inset(att_summary)),
      aes(npcx = npcx, npcy = npcy, label = label),
      hjust = 0, vjust = 1, size = size, color = "black",
      label.size = 0.3, fill = scales::alpha("white", 0.85),
      label.padding = unit(0.3, "lines"), inherit.aes = FALSE
    ),
    coord_cartesian(clip = "off"),
    theme(plot.margin = margin(t = 5, r = margin_r, b = 5, l = 5),
          legend.justification = c("left", "top"))
  )
}

# Bracket annotations on violin/box panels (Fig 1 B&D, Fig 2 B&D, Supp S2/S3 B&D)
# Uses pre-computed Tukey-adjusted p-values from emmeans `pairs()` rather than
# letting ggsignif run its own (unadjusted) test. step_increase auto-stacks
# brackets vertically; expand the panel ylim to leave headroom.
fmt_p_label <- function(p) {
  ifelse(p < 0.001, "p < 0.001", sprintf("p = %.3f", p))
}

# Inset for GAMM panels (Supp Fig S2/S3 A&C): point-estimate amplitudes (no CI)
# + 2-column compact Aug | Dec p-values from emmeans_contrasts_lpmatrix output.
# Headers in UPPERCASE for visual distinction.
make_gamm_inset <- function(amp_df, contrasts_df, by_var, group_levels = NULL) {
  if (!is.null(group_levels)) {
    amp_df <- amp_df[order(match(amp_df[[by_var]], group_levels)), ]
  }
  amp_lines <- sprintf("%s: %.2f", amp_df[[by_var]], amp_df$amplitude)

  cw <- contrasts_df %>%
    select(contrast, MONTH_NUM, p.value) %>%
    tidyr::pivot_wider(names_from = MONTH_NUM, values_from = p.value)

  fmt_p_compact <- function(p) ifelse(p < 0.001, "<.001", sprintf(".%03d", round(p * 1000)))
  pval_lines <- sprintf("%s: %s | %s", cw$contrast,
                        fmt_p_compact(cw$`8`), fmt_p_compact(cw$`12`))

  paste(c("AMPLITUDE (GAMM)", amp_lines, "",
          "AUG | DEC P-VALUES", pval_lines), collapse = "\n")
}

gamm_inset_layer <- function(amp_df, contrasts_df, by_var, group_levels = NULL,
                              npcx = 1.02, npcy = 0.45, size = 2.7,
                              margin_r = 110) {
  list(
    ggpp::geom_label_npc(
      data = tibble(npcx = npcx, npcy = npcy,
                    label = make_gamm_inset(amp_df, contrasts_df, by_var, group_levels)),
      aes(npcx = npcx, npcy = npcy, label = label),
      hjust = 0, vjust = 1, size = size, color = "black",
      label.size = 0.3, fill = scales::alpha("white", 0.85),
      label.padding = unit(0.3, "lines"), inherit.aes = FALSE
    ),
    coord_cartesian(clip = "off"),
    theme(plot.margin = margin(t = 5, r = margin_r, b = 5, l = 5),
          legend.justification = c("left", "top"))
  )
}

# Aug→Dec within-group brackets + interaction-p subtitle for Supp S2/S3 B&D.
# Reads Holm-adjusted contrasts from emmeans(model, pairwise ~ MONTH_LABEL | by_var).
# Bracket spans the two dodged box positions for each group (dodge_width/4 from
# center). Interaction p comes from anova(model) at the by_var:MONTH_LABEL row
# (lmerTest gives Satterthwaite-df F-test p-values).
aug_dec_brackets_layer <- function(model, data, y_var, by_var, group_levels,
                                   interaction_term, by_label,
                                   dodge_width = 0.75, textsize = 3) {
  emm_within <- emmeans::emmeans(model,
                                 as.formula(paste("pairwise ~ MONTH_LABEL |", by_var)),
                                 adjust = "holm")
  cw <- as.data.frame(emm_within$contrasts)
  cw <- cw[match(group_levels, cw[[by_var]]), ]

  n_groups <- length(group_levels)
  half_dodge <- dodge_width / 4
  xmin <- seq_len(n_groups) - half_dodge
  xmax <- seq_len(n_groups) + half_dodge

  # 99th percentile keeps brackets just above the whiskers even when extreme
  # outliers sit outside coord_cartesian's ylim (which would clip the brackets).
  y_max <- quantile(data[[y_var]], 0.99, na.rm = TRUE)
  y_pos <- rep(y_max * 1.04, n_groups)

  int_p <- anova(model)[interaction_term, "Pr(>F)"]
  int_label <- sprintf("%s × Month: %s", by_label, fmt_p_label(int_p))

  list(
    ggsignif::geom_signif(
      annotations = fmt_p_label(cw$p.value),
      xmin = xmin, xmax = xmax,
      y_position = y_pos,
      tip_length = 0.01, textsize = textsize, vjust = -0.2
    ),
    labs(subtitle = int_label),
    theme(plot.subtitle = element_text(face = "bold", color = "black", hjust = 0.5))
  )
}

# One-line β annotation for Figure S5 (geographic daylight regression).
# Sources values from daylight_results_cluster (cluster-robust SEs by state).
beta_inset_layer <- function(daylight_df, outcome, unit,
                             npcx = 1.02, npcy = 0.35, size = 3,
                             margin_r = 130) {
  r <- daylight_df[daylight_df$Outcome == outcome, , drop = FALSE]
  p_str <- if (r$p_value < 0.001) "< 0.001" else sprintf("= %.3f", r$p_value)
  label <- sprintf("β = %.3f %s (95%% CI: %.3f, %.3f); p %s",
                   r$beta_daylight, unit, r$ci_low, r$ci_high, p_str)
  list(
    ggpp::geom_label_npc(
      data = tibble(npcx = npcx, npcy = npcy, label = label),
      aes(npcx = npcx, npcy = npcy, label = label),
      hjust = 0, vjust = 1, size = size, color = "black",
      label.size = 0.3, fill = scales::alpha("white", 0.85),
      label.padding = unit(0.3, "lines"), inherit.aes = FALSE
    ),
    coord_cartesian(clip = "off"),
    theme(plot.margin = margin(t = 5, r = margin_r, b = 5, l = 5),
          legend.justification = c("left", "top"))
  )
}

violin_brackets_layer <- function(tukey_obj, comparisons, group_levels,
                                  y_positions, tip_length = 0.4,
                                  textsize = 3, linewidth = 0.4) {
  # Manual brackets via geom_segment + geom_text. Replaces a prior ggsignif
  # implementation (broken under ggplot2 4.x as of ggsignif 0.6.4). Caller
  # supplies group_levels (factor order on x-axis) and y_positions (one per
  # comparison, in data units of the y-axis).
  td <- as.data.frame(tukey_obj)
  contrasts_str <- as.character(td$contrast)
  pair_keys <- vapply(strsplit(contrasts_str, " - ", fixed = TRUE),
                      function(x) paste(sort(x), collapse = "|"),
                      character(1))
  p_lookup <- setNames(td$p.value, pair_keys)

  brackets <- do.call(rbind, lapply(seq_along(comparisons), function(i) {
    comp <- comparisons[[i]]
    key <- paste(sort(comp), collapse = "|")
    p <- p_lookup[[key]]
    label <- if (is.null(p) || is.na(p)) "p = ?" else fmt_p_label(p)
    data.frame(
      x1 = match(comp[1], group_levels),
      x2 = match(comp[2], group_levels),
      y  = y_positions[i],
      label = label,
      stringsAsFactors = FALSE
    )
  }))

  list(
    # Horizontal bracket bar
    geom_segment(data = brackets,
                 aes(x = x1, xend = x2, y = y, yend = y),
                 linewidth = linewidth, inherit.aes = FALSE),
    # Left tip (down-pointing)
    geom_segment(data = brackets,
                 aes(x = x1, xend = x1, y = y - tip_length, yend = y),
                 linewidth = linewidth, inherit.aes = FALSE),
    # Right tip (down-pointing)
    geom_segment(data = brackets,
                 aes(x = x2, xend = x2, y = y - tip_length, yend = y),
                 linewidth = linewidth, inherit.aes = FALSE),
    # Label above bar
    geom_text(data = brackets,
              aes(x = (x1 + x2) / 2, y = y + tip_length * 0.5, label = label),
              size = textsize, vjust = 0, inherit.aes = FALSE)
  )
}
```

### Make monthly dataframe
```{r}
df <- df %>%
  mutate(
    LOWER_DAYS = as.Date(LOWER_DAYS),
    MONTH = floor_date(LOWER_DAYS, "month")
  )

df_month <- df %>%
  group_by(USER_ID, MONTH) %>%
  summarise(
    RESTING_HEART_RATE = mean(RESTING_HEART_RATE, na.rm = TRUE),
    HRV = mean(HRV, na.rm = TRUE),
    SHRZS = mean(SHRZS, na.rm = TRUE),
    MINUTES_OF_SLEEP = mean(MINUTES_OF_SLEEP, na.rm = TRUE),

    # Static user fields
    STRAP_GENERATION = first(na.omit(STRAP_GENERATION)),
    HEIGHT = first(na.omit(HEIGHT)),
    WEIGHT = first(na.omit(WEIGHT)),
    GENDER = first(na.omit(GENDER)),
    BMI = first(na.omit(BMI)),
    AGE_2023 = first(na.omit(AGE_2023)),
    AGE_BUCKET = first(na.omit(AGE_BUCKET)),
    STATE = first(na.omit(STATE)),
    .groups = 'drop'
  )

df_month <- df_month %>%
  group_by(USER_ID) %>%
  mutate(
    # Person-level mean (between-person)
    SHRZS_b = mean(SHRZS, na.rm = TRUE),
    MINUTES_OF_SLEEP_b = mean(MINUTES_OF_SLEEP, na.rm = TRUE),

    # Person-mean centered (within-person)
    SHRZS_w = SHRZS - SHRZS_b,
    MINUTES_OF_SLEEP_w = MINUTES_OF_SLEEP - MINUTES_OF_SLEEP_b
  ) %>%
  ungroup()
```

### Compute some descriptive stats
```{r}
# Compute some descriptive stats
df_person <- df_month %>%
  group_by(USER_ID) %>%
  summarise(
    RESTING_HEART_RATE = mean(RESTING_HEART_RATE, na.rm = TRUE),
    HRV = mean(HRV, na.rm = TRUE),
    SHRZS = mean(SHRZS, na.rm = TRUE),
    MINUTES_OF_SLEEP = mean(MINUTES_OF_SLEEP, na.rm = TRUE),

    # Retain static attributes
    BMI = first(na.omit(BMI)),
    AGE_2023 = first(na.omit(AGE_2023)),
    GENDER = first(na.omit(GENDER)),
    AGE_BUCKET = first(na.omit(AGE_BUCKET)),
    .groups = "drop"
  )

describe_vars <- function(data, group_name = "Overall") {
  vars <- c("RESTING_HEART_RATE", "HRV", "SHRZS", "MINUTES_OF_SLEEP", "BMI", "AGE_2023")
  
  summary_stats <- data %>%
    summarise(across(all_of(vars),
                     list(mean = ~mean(.x, na.rm = TRUE),
                          sd = ~sd(.x, na.rm = TRUE)),
                     .names = "{.col}_{.fn}"))
  
  summary_stats$Group <- group_name
  return(summary_stats)
}

# Overall
overall_stats <- describe_vars(df_person, "Overall")

# By GENDER
gender_stats <- df_person %>%
  group_by(GENDER) %>%
  group_modify(~describe_vars(.x, unique(.x$GENDER)))

# By AGE_BUCKET
agebucket_stats <- df_person %>%
  group_by(AGE_BUCKET) %>%
  group_modify(~describe_vars(.x, unique(.x$AGE_BUCKET)))

# Combine all results
descriptive_summary <- bind_rows(overall_stats, gender_stats, agebucket_stats)

# Variables to test
vars <- c(
  "RESTING_HEART_RATE",
  "HRV",
  "SHRZS",
  "MINUTES_OF_SLEEP",
  "BMI",
  "AGE_2023"
)

# Run Welch t-tests for each variable
gender_ttests <- map_df(vars, function(v) {

  tt <- t.test(df_person[[v]] ~ df_person$GENDER)

  tibble(
    Variable = v,
    mean_male = mean(df_person[[v]][df_person$GENDER == "male"], na.rm = TRUE),
    mean_female = mean(df_person[[v]][df_person$GENDER == "female"], na.rm = TRUE),
    diff = mean_female - mean_male,
    t = as.numeric(tt$statistic),
    p = round(tt$p.value, 4)
  )
})

# Apply Holm correction across all variables
gender_ttests$p_holm <- round(p.adjust(gender_ttests$p, method = "holm"), 4)

gender_ttests

# Run ANOVA for each variable
agebucket_anova <- map_df(vars, function(v) {

  model <- aov(df_person[[v]] ~ df_person$AGE_BUCKET)
  tidy_res <- tidy(model)[1, ]  # Extract only the AGE_BUCKET row

  tibble(
    Variable = v,
    F = round(tidy_res$statistic, 2),
    p = round(tidy_res$p.value, 4)
  )
})

# Apply Holm correction across all variables
agebucket_anova$p_holm <- round(p.adjust(agebucket_anova$p, method = "holm"), 4)

agebucket_anova

# Variables to follow up on (excluding BMI)
followup_vars <- c(
  "RESTING_HEART_RATE",
  "HRV",
  "SHRZS",
  "MINUTES_OF_SLEEP",
  "AGE_2023"
)

# Run ANOVA + Tukey HSD per variable
tukey_results <- map_df(followup_vars, function(v) {
  
  model <- aov(df_person[[v]] ~ df_person$AGE_BUCKET)
  tk <- TukeyHSD(model, "df_person$AGE_BUCKET")
  
  as_tibble(tk$`df_person$AGE_BUCKET`, rownames = "contrast") %>%
    mutate(
      Variable = v,
      estimate = round(diff, 4),
      p_adj = round(`p adj`, 4)
    ) %>%
    select(Variable, contrast, estimate, p_adj)
})

tukey_results

```

### Create field for month and compute RHR/HRV offset (decompose)
```{r}
# Create field for month
df_month <- df_month %>%
  mutate(
    MONTH_NUM = month(MONTH),  
    GENDER = as.factor(GENDER),
    USER_ID = as.factor(USER_ID)
  )

df_month <- df_month %>%
  group_by(USER_ID) %>%
  mutate(
    # Between-person RHR: average per user
    RHR_person_mean = mean(RESTING_HEART_RATE, na.rm = TRUE),
    
    # Within-person RHR: deviation from person's mean (centered RHR)
    RHR_within = RESTING_HEART_RATE - RHR_person_mean
  ) %>%
  ungroup()

df_month <- df_month %>%
  group_by(USER_ID) %>%
  mutate(
    # Between-person HRV: average per user
    HRV_person_mean = mean(HRV, na.rm = TRUE),
    
    # Within-person HRV: deviation from person's mean (centered HRV)
    HRV_within = HRV - HRV_person_mean
  ) %>%
  ungroup()
```


# AIM 1
## RHR Overall
```{r}
# Baseline RHR model (stored for reuse in Aim 2 attenuation analyses)
model_rhr_baseline <- bam(
  RHR_within ~
    s(MONTH_NUM, bs = "cc", k = 12) +
    s(AGE_2023) +
    GENDER +
    BMI +
    RHR_person_mean,
  data = df_month,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

v_df_rhr_overall <- predict_marginal_trajectory(model_rhr_baseline, data = df_month)

p_s1a <- ggplot(v_df_rhr_overall, aes(x = MONTH_NUM, y = visregFit)) +
  geom_line(color = "black", linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr),
              fill = "black", alpha = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "RHR Offset (bpm)") +
  theme_seasonality(legend = FALSE)
p_s1a

# ---- Identify modeled peak/trough and amplitude ----
peak_month <- v_df_rhr_overall %>% filter(visregFit == max(visregFit)) %>% select(MONTH_NUM, visregFit)
trough_month <- v_df_rhr_overall %>% filter(visregFit == min(visregFit)) %>% select(MONTH_NUM, visregFit)
modeled_amplitude <- peak_month$visregFit - trough_month$visregFit

# Bootstrap amplitude (participant-level resampling)
formula_rhr_baseline <- RHR_within ~ s(MONTH_NUM, bs = "cc", k = 12) + s(AGE_2023) + GENDER + BMI + RHR_person_mean

boot_rhr_overall <- bootstrap_amplitude(df_month, formula_rhr_baseline)

amplitude_ci <- quantile(boot_rhr_overall$amplitude, probs = c(0.025, 0.975), na.rm = TRUE)
amplitude_mean <- mean(boot_rhr_overall$amplitude, na.rm = TRUE)

cat("Bootstrapped amplitude mean:", round(amplitude_mean, 3), "\n")
cat("95% CI:", round(amplitude_ci[1], 3), "-", round(amplitude_ci[2], 3), "\n")
```

## HRV Overall
```{r}
# Baseline HRV model (stored for reuse in Aim 2 attenuation analyses)
model_hrv_baseline <- bam(
  HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12) + s(AGE_2023) + GENDER + BMI + HRV_person_mean,
  data = df_month,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

v_df_hrv_overall <- predict_marginal_trajectory(model_hrv_baseline, data = df_month)

p_s1b <- ggplot(v_df_hrv_overall, aes(x = MONTH_NUM, y = visregFit)) +
  geom_line(color = "black", linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), fill = "black", alpha = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "HRV Offset (ms)") +
  theme_seasonality(legend = FALSE)
p_s1b

# ---- Identify modeled peak/trough and amplitude ----
peak_month <- v_df_hrv_overall %>% filter(visregFit == max(visregFit)) %>% select(MONTH_NUM, visregFit)
trough_month <- v_df_hrv_overall %>% filter(visregFit == min(visregFit)) %>% select(MONTH_NUM, visregFit)
modeled_amplitude <- peak_month$visregFit - trough_month$visregFit

# Bootstrap amplitude (participant-level resampling)
formula_hrv_baseline <- HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12) + s(AGE_2023) + GENDER + BMI + HRV_person_mean

boot_hrv_overall <- bootstrap_amplitude(df_month, formula_hrv_baseline)

amplitude_ci <- quantile(boot_hrv_overall$amplitude, probs = c(0.025, 0.975), na.rm = TRUE)
amplitude_mean <- mean(boot_hrv_overall$amplitude, na.rm = TRUE)

cat("Bootstrapped amplitude mean:", round(amplitude_mean, 3), "\n")
cat("95% CI:", round(amplitude_ci[1], 3), "-", round(amplitude_ci[2], 3), "\n")
```


## RHR by AGE
### Model 1 - GAM
```{r}
df_month <- df_month %>%
  mutate(AGE_BUCKET = factor(AGE_BUCKET, levels = c("Young", "Middle", "Old")))

# ---- Fit GAMM model with cyclic spline by AGE_BUCKET ----
model_rhr_age <- bam(
  RHR_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = AGE_BUCKET) + AGE_BUCKET + GENDER + BMI + RHR_person_mean,
  data = df_month,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# Extract smooths via marginal-trajectory helper
v_df <- predict_marginal_trajectory(
  model_rhr_age,
  data = df_month,
  by_var = "AGE_BUCKET"
) %>%
  mutate(AGE_BUCKET = factor(AGE_BUCKET, levels = c("Young", "Middle", "Old")))


# ---- Plot ----
p_fig1a <- ggplot(v_df, aes(x = MONTH_NUM, y = visregFit, color = AGE_BUCKET, fill = AGE_BUCKET)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "Age Group", values = age_colors) +
  scale_fill_manual(name = "Age Group", values = age_colors) +
  scale_x_continuous(
    breaks = 1:12,
    labels = month.abb,
    name = NULL
  ) +
  scale_y_continuous(name = "RHR Offset (bpm)") +
  theme_seasonality()
p_fig1a


# ---- Bootstrap modeled amplitude by AGE_BUCKET ----
formula_rhr_age <- RHR_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = AGE_BUCKET) + AGE_BUCKET + GENDER + BMI + RHR_person_mean

boot_rhr_age <- bootstrap_amplitude(df_month, formula_rhr_age, by = "AGE_BUCKET")

bootstrap_summary <- boot_rhr_age %>%
  group_by(AGE_BUCKET) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

boot_pvals <- bootstrap_pairwise(boot_rhr_age, "AGE_BUCKET")
```
### Model 2 - GAMM
```{r}
set.seed(2024)

subsample_ids <- df_month %>%
  distinct(USER_ID, AGE_BUCKET, GENDER) %>%
  group_by(AGE_BUCKET, GENDER) %>%
  slice_sample(n = 1000) %>%
  ungroup()

df_sub <- df_month %>%
  semi_join(subsample_ids, by = "USER_ID")

df_sub <- df_sub %>%
  mutate(USER_ID = as.factor(USER_ID))

df_sub <- df_sub %>%
  mutate(AGE_BUCKET = factor(AGE_BUCKET, levels = c("Young", "Middle", "Old")))

model_rhr_age_mixed <- bam(
  RHR_within ~ 
    s(MONTH_NUM, bs = "cc", k = 12, by = AGE_BUCKET) + 
    AGE_BUCKET + GENDER + BMI + RHR_person_mean + 
    s(USER_ID, bs = "re"),
  data = df_sub,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

summary(model_rhr_age_mixed)

# Extract smooths via marginal-trajectory helper (excluding random effect)
v_mixed_df <- predict_marginal_trajectory(
  model_rhr_age_mixed,
  data = df_sub,
  by_var = "AGE_BUCKET",
  exclude_re = "s(USER_ID)"
) %>%
  mutate(AGE_BUCKET = factor(AGE_BUCKET, levels = c("Young", "Middle", "Old")))


# Plot
p_s2a <- ggplot(v_mixed_df, aes(x = MONTH_NUM, y = visregFit, color = AGE_BUCKET, fill = AGE_BUCKET)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "Age Group", values = age_colors) +
  scale_fill_manual(name = "Age Group", values = age_colors) +
  scale_x_continuous(
    breaks = 1:12,
    labels = month.abb,
    name = NULL
  ) +
  scale_y_continuous(name = "RHR Offset (bpm)") +
  theme_seasonality()
p_s2a

#Extract amplitude values
amplitude_gamm <- v_mixed_df %>%
  group_by(AGE_BUCKET) %>%
  summarise(
    amplitude = max(visregFit, na.rm = TRUE) - min(visregFit, na.rm = TRUE),
    peak_month = MONTH_NUM[which.max(visregFit)],
    trough_month = MONTH_NUM[which.min(visregFit)],
    .groups = "drop"
  )

## EMMEANS Comparisons via lpmatrix
res <- emmeans_contrasts_lpmatrix(model_rhr_age_mixed, data = df_sub, by_var = "AGE_BUCKET")
emm_manual    <- res$emm
contrasts_all <- res$contrasts

print("Estimated Marginal Means:")
print(emm_manual)
print("Pairwise Contrasts:")
print(contrasts_all)
```
### Model 3 - Individual Level Amplitude
```{r}
# ---- Compute per-person RHR amplitude: December - August ----
df_amplitude <- df_month %>%
  filter(MONTH_NUM %in% c(8, 12)) %>%
  mutate(SEASON = case_when(
    MONTH_NUM == 8 ~ "August",
    MONTH_NUM == 12 ~ "December"
  )) %>%
  group_by(USER_ID, SEASON) %>%
  summarise(
    mean_rhr = mean(RHR_within, na.rm = TRUE),
    AGE_BUCKET = first(AGE_BUCKET),
    AGE_2023 = first(AGE_2023),
    BMI = first(BMI),
    GENDER = first(GENDER),
    RHR_person_mean = first(RHR_person_mean),
    .groups = "drop"
  ) %>%
  pivot_wider(names_from = SEASON, values_from = mean_rhr) %>%
  mutate(
    amplitude = December - August  # RHR: winter minus summer
  )

# ---- Per-person amplitude summary by age bucket ----
amplitude_summary <- df_amplitude %>%
  group_by(AGE_BUCKET) %>%
  summarise(
    mean_amplitude = mean(amplitude, na.rm = TRUE),
    sd_amplitude = sd(amplitude, na.rm = TRUE),
    n = n(),
    mean_sd = paste0(round(mean_amplitude, 2), " ± ", round(sd_amplitude, 2)),
    .groups = "drop"
  )


# ---- Linear model predicting per-person amplitude ----
model_amplitude <- lm(amplitude ~ AGE_BUCKET + GENDER + BMI + RHR_person_mean, data = df_amplitude)
anova(model_amplitude)
# ---- 95% Confidence Intervals for model coefficients ----
confint(model_amplitude)

# Get estimated marginal means
emm <- emmeans(model_amplitude, ~ AGE_BUCKET)
# Pairwise Tukey comparisons
tukey_results <- pairs(emm, adjust = "tukey")
# View results
summary(tukey_results, infer = c(TRUE, TRUE))  # Adds CIs and p-values


df_amplitude <- df_amplitude %>%
  mutate(AGE_BUCKET = factor(AGE_BUCKET, levels = c("Young", "Middle", "Old")))

p_fig1b <- ggplot(df_amplitude, aes(x = AGE_BUCKET, y = amplitude, fill = AGE_BUCKET)) +
  geom_violin(trim = FALSE, alpha = 0.5, color = NA) +
  geom_boxplot(width = 0.1, outlier.shape = NA, alpha = 0.7) +
  stat_summary(fun = mean, geom = "point", shape = 21, size = 3, fill = "white") +
  scale_fill_manual(values = age_colors) +
  scale_y_continuous(name = "RHR Amplitude (Dec − Aug, bpm)") +
  coord_cartesian(ylim = c(-5, 13)) +
  labs(x = NULL, y = "RHR Amplitude (Dec − Aug, bpm)") +
  theme_seasonality(legend = FALSE) +
  violin_brackets_layer(tukey_results,
    comparisons = list(c("Young", "Middle"), c("Middle", "Old"), c("Young", "Old")),
    group_levels = c("Young", "Middle", "Old"),
    y_positions = c(10.5, 11.7, 12.9))
p_fig1b
```
### Model 4 - Box and Whisker of Raw RHR Values in Peak and Trough Months by Age
```{r}
# Filter for August and December
df_aug_dec <- df_month %>%
  filter(MONTH_NUM %in% c(8, 12)) %>%
  mutate(
    MONTH_LABEL = factor(ifelse(MONTH_NUM == 8, "August", "December"), levels = c("August", "December")),
    AGE_BUCKET = factor(AGE_BUCKET, levels = c("Young", "Middle", "Old")),
    pattern = ifelse(MONTH_LABEL == "December", "stripe", "none")
  )
# Plot using ggpattern
p_s2b <- ggplot(df_aug_dec, aes(x = AGE_BUCKET, y = RESTING_HEART_RATE)) +
  geom_boxplot_pattern(
    aes(
      fill = AGE_BUCKET,
      pattern = MONTH_LABEL,
      pattern_fill = AGE_BUCKET
    ),
    position = position_dodge(width = 0.75),
    color = "black",
    pattern_density = 0.4,
    pattern_spacing = 0.05,
    alpha = 0.7,
    outlier.shape = NA
  ) +
  stat_summary(
    fun = median,
    geom = "point",
    aes(group = interaction(MONTH_LABEL, AGE_BUCKET)),
    position = position_dodge(width = 0.75),
    shape = 21,
    size = 3,
    fill = "white",
    color = "black"
  ) +
  scale_fill_manual(values = age_colors) +
  scale_pattern_fill_manual(values = age_colors) +
  scale_pattern_manual(
    values = c("August" = "none", "December" = "stripe"),
    name = "Month"
  ) +
  guides(
    fill = "none",
    pattern_fill = "none",
    pattern = guide_legend(
      override.aes = list(
        fill = "grey70",
        pattern_density = 0.5,
        pattern_spacing = 0.02
      )
    )
  ) +
  labs(
    x = NULL,
    y = "Resting Heart Rate (bpm)"
  ) +
  theme_seasonality()
p_s2b

# Fit linear mixed model with AGE_BUCKET * MONTH_NUM interaction
model_aug_dec_rhr_age <- lmer(
  RESTING_HEART_RATE ~ AGE_BUCKET * MONTH_LABEL + GENDER + BMI + (1 | USER_ID),
  data = df_aug_dec
)

anova(model_aug_dec_rhr_age)

# Age group comparisons within each month
emmeans(model_aug_dec_rhr_age, pairwise ~ AGE_BUCKET | MONTH_LABEL, adjust = "holm")

# August vs December within each age group
emmeans(model_aug_dec_rhr_age, pairwise ~ MONTH_LABEL | AGE_BUCKET, adjust = "holm")

# Add within-group Aug→Dec brackets + interaction-p subtitle
p_s2b <- p_s2b + aug_dec_brackets_layer(
  model_aug_dec_rhr_age, df_aug_dec, y_var = "RESTING_HEART_RATE",
  by_var = "AGE_BUCKET", group_levels = c("Young", "Middle", "Old"),
  interaction_term = "AGE_BUCKET:MONTH_LABEL", by_label = "Age"
)
```


## HRV by AGE
### Model 1 - GAM
```{r}
# GAM model for HRV
model_hrv_age <- bam(
  HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = AGE_BUCKET) + AGE_BUCKET + GENDER + BMI + HRV_person_mean,
  data = df_month,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)


v_hrv_df <- predict_marginal_trajectory(
  model_hrv_age,
  data = df_month,
  by_var = "AGE_BUCKET"
) %>%
  mutate(AGE_BUCKET = factor(AGE_BUCKET, levels = c("Young", "Middle", "Old")))

p_fig1c <- ggplot(v_hrv_df, aes(x = MONTH_NUM, y = visregFit, color = AGE_BUCKET, fill = AGE_BUCKET)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "Age Group", values = age_colors) +
  scale_fill_manual(name = "Age Group", values = age_colors) +
  scale_x_continuous(
    breaks = 1:12,
    labels = month.abb,
    name = NULL
  ) +
  scale_y_continuous(name = "HRV Offset (ms)") +
  theme_seasonality()
p_fig1c


# ---- Bootstrap modeled amplitude by AGE_BUCKET ----
formula_hrv_age <- HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = AGE_BUCKET) + AGE_BUCKET + GENDER + BMI + HRV_person_mean

boot_hrv_age <- bootstrap_amplitude(df_month, formula_hrv_age, by = "AGE_BUCKET")

bootstrap_summary_hrv_age <- boot_hrv_age %>%
  group_by(AGE_BUCKET) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

boot_pvals_hrv_age <- bootstrap_pairwise(boot_hrv_age, "AGE_BUCKET")
```
### Model 2 - GAMM
```{r}
model_hrv_age_mixed <- bam(
  HRV_within ~ 
    s(MONTH_NUM, bs = "cc", k = 12, by = AGE_BUCKET) +
    AGE_BUCKET + GENDER + BMI + HRV_person_mean +
    s(USER_ID, bs = "re"),
  data = df_sub,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

summary(model_hrv_age_mixed)


v_mixed_hrv_df <- predict_marginal_trajectory(
  model_hrv_age_mixed,
  data = df_sub,
  by_var = "AGE_BUCKET",
  exclude_re = "s(USER_ID)"
) %>%
  mutate(AGE_BUCKET = factor(AGE_BUCKET, levels = c("Young", "Middle", "Old")))

p_s2c <- ggplot(v_mixed_hrv_df, aes(x = MONTH_NUM, y = visregFit, color = AGE_BUCKET, fill = AGE_BUCKET)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "Age Group", values = age_colors) +
  scale_fill_manual(name = "Age Group", values = age_colors) +
  scale_x_continuous(
    breaks = 1:12,
    labels = month.abb,
    name = NULL
  ) +
  scale_y_continuous(name = "HRV Offset (ms)") +
  theme_seasonality()
p_s2c

amplitude_gamm_hrv <- v_mixed_hrv_df %>%
  group_by(AGE_BUCKET) %>%
  summarise(
    amplitude = max(visregFit, na.rm = TRUE) - min(visregFit, na.rm = TRUE),
    peak_month = MONTH_NUM[which.max(visregFit)],
    trough_month = MONTH_NUM[which.min(visregFit)],
    .groups = "drop"
  )

# --- HRV EMMEANS comparisons via lpmatrix ---
res_hrv <- emmeans_contrasts_lpmatrix(model_hrv_age_mixed, data = df_sub, by_var = "AGE_BUCKET")
emm_manual_hrv    <- res_hrv$emm
contrasts_all_hrv <- res_hrv$contrasts

print("Estimated Marginal Means (HRV):")
print(emm_manual_hrv)
print("Pairwise Contrasts (HRV):")
print(contrasts_all_hrv)
```
### Model 3 - Invividual Level Amplitude
```{r}
df_amplitude_hrv <- df_month %>%
  filter(MONTH_NUM %in% c(8, 12)) %>%
  mutate(SEASON = case_when(
    MONTH_NUM == 8 ~ "August",
    MONTH_NUM == 12 ~ "December"
  )) %>%
  group_by(USER_ID, SEASON) %>%
  summarise(
    mean_hrv = mean(HRV_within, na.rm = TRUE),
    AGE_BUCKET = first(AGE_BUCKET),
    AGE_2023 = first(AGE_2023),
    BMI = first(BMI),
    GENDER = first(GENDER),
    HRV_person_mean = first(HRV_person_mean),
    .groups = "drop"
  ) %>%
  pivot_wider(names_from = SEASON, values_from = mean_hrv) %>%
  mutate(
    amplitude = August - December   
  )

amplitude_summary_hrv <- df_amplitude_hrv %>%
  group_by(AGE_BUCKET) %>%
  summarise(
    mean_amplitude = mean(amplitude, na.rm = TRUE),
    sd_amplitude = sd(amplitude, na.rm = TRUE),
    n = n(),
    mean_sd = paste0(round(mean_amplitude, 2), " ± ", round(sd_amplitude, 2)),
    .groups = "drop"
  )

model_amplitude_hrv <- lm(amplitude ~ AGE_BUCKET + GENDER + BMI + HRV_person_mean, 
                          data = df_amplitude_hrv)

anova(model_amplitude_hrv)
confint(model_amplitude_hrv)

emm_hrv <- emmeans(model_amplitude_hrv, ~ AGE_BUCKET)

tukey_results_hrv <- pairs(emm_hrv, adjust = "tukey")

summary(tukey_results_hrv, infer = c(TRUE, TRUE))

df_amplitude_hrv <- df_amplitude_hrv %>%
  mutate(AGE_BUCKET = factor(AGE_BUCKET, levels = c("Young", "Middle", "Old")))

p_fig1d <- ggplot(df_amplitude_hrv, aes(x = AGE_BUCKET, y = amplitude, fill = AGE_BUCKET)) +
  geom_violin(trim = FALSE, alpha = 0.5, color = NA) +
  geom_boxplot(width = 0.1, outlier.shape = NA, alpha = 0.7) +
  stat_summary(fun = mean, geom = "point", shape = 21, size = 3, fill = "white") +
  scale_fill_manual(values = age_colors) +
  scale_y_continuous(name = "HRV Amplitude (Aug − Dec, ms)") +
  coord_cartesian(ylim = c(-20, 40)) +
  labs(x = NULL, y = "HRV Amplitude (Aug − Dec, ms)") +
  theme_seasonality(legend = FALSE) +
  violin_brackets_layer(tukey_results_hrv,
    comparisons = list(c("Young", "Middle"), c("Middle", "Old"), c("Young", "Old")),
    group_levels = c("Young", "Middle", "Old"),
    y_positions = c(31, 34, 37),
    tip_length = 1)
p_fig1d
```
### Model 4 - Box and Whisker of Raw HRV values in Peak and Trough Months by Age
```{r}
# Plot
p_s2d <- ggplot(df_aug_dec, aes(x = AGE_BUCKET, y = HRV)) +
  geom_boxplot_pattern(
    aes(
      fill = AGE_BUCKET,
      pattern = MONTH_LABEL,
      pattern_fill = AGE_BUCKET
    ),
    position = position_dodge(width = 0.75),
    color = "black",
    pattern_density = 0.4,
    pattern_spacing = 0.05,
    alpha = 0.7,
    outlier.shape = NA
  ) +
  stat_summary(
    fun = median,
    geom = "point",
    aes(group = interaction(MONTH_LABEL, AGE_BUCKET)),
    position = position_dodge(width = 0.75),
    shape = 21,
    size = 3,
    fill = "white",
    color = "black"
  ) +
  scale_fill_manual(values = age_colors) +
  scale_pattern_fill_manual(values = age_colors) +
  scale_pattern_manual(
    values = c("August" = "none", "December" = "stripe"),
    name = "Month"
  ) +
  guides(
    fill = "none",
    pattern_fill = "none",
    pattern = guide_legend(
      override.aes = list(
        fill = "grey70",
        pattern_density = 0.5,
        pattern_spacing = 0.02
      )
    )
  ) +
  labs(
    x = NULL,
    y = "Heart Rate Variability (ms)"
  ) +
  coord_cartesian(ylim = c(0, 150)) +
  theme_seasonality()
p_s2d

# ---- Step 4: Linear Mixed Model ----
model_aug_dec_hrv_age <- lmer(
  HRV ~ AGE_BUCKET * MONTH_LABEL + GENDER + BMI + (1 | USER_ID),
  data = df_aug_dec
)

# ---- Step 5: ANOVA table ----
anova(model_aug_dec_hrv_age)

# ---- Step 6: Pairwise comparisons ----

# Age group comparisons within each month
emmeans(model_aug_dec_hrv_age, pairwise ~ AGE_BUCKET | MONTH_LABEL, adjust = "holm")

# August vs December within each age group
emmeans(model_aug_dec_hrv_age, pairwise ~ MONTH_LABEL | AGE_BUCKET, adjust = "holm")

# Add within-group Aug→Dec brackets + interaction-p subtitle
p_s2d <- p_s2d + aug_dec_brackets_layer(
  model_aug_dec_hrv_age, df_aug_dec, y_var = "HRV",
  by_var = "AGE_BUCKET", group_levels = c("Young", "Middle", "Old"),
  interaction_term = "AGE_BUCKET:MONTH_LABEL", by_label = "Age"
)
```


## RHR by SEX
### Model 1 - GAM
```{r}
df_month <- df_month %>%
  mutate(GENDER = factor(GENDER, levels = c("male", "female")))  

model_rhr_gender <- bam(
  RHR_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = GENDER) + GENDER + s(AGE_2023) + BMI + RHR_person_mean,
  data = df_month,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

v_df_gender <- predict_marginal_trajectory(
  model_rhr_gender,
  data = df_month,
  by_var = "GENDER"
) %>%
  mutate(GENDER = factor(GENDER, levels = c("male", "female")))

p_fig2a <- ggplot(v_df_gender, aes(x = MONTH_NUM, y = visregFit, color = GENDER, fill = GENDER)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "Sex", values = gender_colors, labels = gender_labels) +
  scale_fill_manual(name = "Sex", values = gender_colors, labels = gender_labels) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "RHR Offset (bpm)") +
  theme_seasonality()

p_fig2a

formula_rhr_sex <- RHR_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = GENDER) + GENDER + s(AGE_2023) + BMI + RHR_person_mean

boot_rhr_sex <- bootstrap_amplitude(df_month, formula_rhr_sex, by = "GENDER")

bootstrap_summary_sex_rhr <- boot_rhr_sex %>%
  group_by(GENDER) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

boot_pvals_sex_rhr <- bootstrap_pairwise(boot_rhr_sex, "GENDER")
```
### Model 2 - GAMM
```{r}
model_rhr_gender_mixed <- bam(
  RHR_within ~ 
    s(MONTH_NUM, bs = "cc", k = 12, by = GENDER) +
    GENDER + s(AGE_2023) + BMI + RHR_person_mean +
    s(USER_ID, bs = "re"),
  data = df_sub,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

v_mixed_gender_df <- predict_marginal_trajectory(
  model_rhr_gender_mixed,
  data = df_sub,
  by_var = "GENDER",
  exclude_re = "s(USER_ID)"
) %>%
  mutate(GENDER = factor(GENDER, levels = c("male", "female")))

p_s3a <- ggplot(v_mixed_gender_df, aes(x = MONTH_NUM, y = visregFit, color = GENDER, fill = GENDER)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "Sex", values = gender_colors, labels = gender_labels) +
  scale_fill_manual(name = "Sex", values = gender_colors, labels = gender_labels) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "RHR Offset (bpm)") +
  theme_seasonality()

p_s3a

amplitude_gamm_gender <- v_mixed_gender_df %>%
  group_by(GENDER) %>%
  summarise(
    amplitude = max(visregFit, na.rm = TRUE) - min(visregFit, na.rm = TRUE),
    peak_month = MONTH_NUM[which.max(visregFit)],
    trough_month = MONTH_NUM[which.min(visregFit)],
    .groups = "drop"
  )

# --- RHR-by-Gender EMMEANS comparisons via lpmatrix ---
res_gender <- emmeans_contrasts_lpmatrix(model_rhr_gender_mixed, data = df_sub, by_var = "GENDER")
emm_manual_gender <- res_gender$emm
gender_contrasts  <- res_gender$contrasts

print("Estimated Marginal Means (RHR by Gender):")
print(emm_manual_gender)
print("Gender Contrasts (RHR):")
print(gender_contrasts)
```
### Model 3 - Individual Level Amplitude
```{r}
df_amplitude_gender <- df_month %>%
  filter(MONTH_NUM %in% c(8, 12)) %>%
  mutate(SEASON = case_when(
    MONTH_NUM == 8 ~ "August",
    MONTH_NUM == 12 ~ "December"
  )) %>%
  group_by(USER_ID, SEASON) %>%
  summarise(
    mean_rhr = mean(RHR_within, na.rm = TRUE),
    GENDER = first(GENDER),
    AGE_2023 = first(AGE_2023),
    BMI = first(BMI),
    RHR_person_mean = first(RHR_person_mean),
    .groups = "drop"
  ) %>%
  pivot_wider(names_from = SEASON, values_from = mean_rhr) %>%
  mutate(
    amplitude = December - August
  )

amplitude_summary_gender <- df_amplitude_gender %>%
  group_by(GENDER) %>%
  summarise(
    mean_amplitude = mean(amplitude, na.rm = TRUE),
    sd_amplitude = sd(amplitude, na.rm = TRUE),
    n = n(),
    mean_sd = paste0(round(mean_amplitude, 2), " ± ", round(sd_amplitude, 2)),
    .groups = "drop"
  )

model_amplitude_gender <- lm(amplitude ~ GENDER + splines::ns(AGE_2023, df = 3) + BMI + RHR_person_mean, data = df_amplitude_gender)
anova(model_amplitude_gender)

confint(model_amplitude_gender)

emm_gender <- emmeans(model_amplitude_gender, ~ GENDER)
tukey_gender <- pairs(emm_gender, adjust = "tukey")
summary(tukey_gender, infer = c(TRUE, TRUE))

# Set order of GENDER for plotting
df_amplitude_gender <- df_amplitude_gender %>%
  mutate(GENDER = factor(GENDER, levels = c("male", "female")))

p_fig2b <- ggplot(df_amplitude_gender, aes(x = GENDER, y = amplitude, fill = GENDER)) +
  geom_violin(trim = FALSE, alpha = 0.5, color = NA) +
  geom_boxplot(width = 0.1, outlier.shape = NA, alpha = 0.7) +
  stat_summary(fun = mean, geom = "point", shape = 21, size = 3, fill = "white") +
  scale_fill_manual(values = gender_colors) +
  scale_x_discrete(labels = gender_labels) +
  scale_y_continuous(name = "RHR Amplitude (Dec - Aug, bpm)") +
  coord_cartesian(ylim = c(-5, 12)) +
  labs(x = NULL, y = "RHR Amplitude (Dec - Aug, bpm)") +
  theme_seasonality(legend = FALSE) +
  violin_brackets_layer(tukey_gender,
    comparisons = list(c("male", "female")),
    group_levels = c("male", "female"),
    y_positions = c(10.5))

p_fig2b

```
### Model 4 - Box and Whisker of Raw RHR Values in Peak and Trough Months by Sex
```{r}
# ---- Boxplot of Raw RHR ----
# Order x-axis male, female to match panel D (Figure S3D)
df_aug_dec <- df_aug_dec %>%
  mutate(GENDER = factor(GENDER, levels = c("male", "female")))
p_s3b <- ggplot(df_aug_dec, aes(x = GENDER, y = RESTING_HEART_RATE)) +
  geom_boxplot_pattern(
    aes(
      fill = GENDER,
      pattern = MONTH_LABEL,
      pattern_fill = GENDER
    ),
    position = position_dodge(width = 0.75),
    color = "black",
    pattern_density = 0.4,
    pattern_spacing = 0.05,
    alpha = 0.7,
    outlier.shape = NA
  ) +
  stat_summary(
    fun = median,
    geom = "point",
    aes(group = interaction(MONTH_LABEL, GENDER)),
    position = position_dodge(width = 0.75),
    shape = 21,
    size = 3,
    fill = "white",
    color = "black"
  ) +
  scale_fill_manual(values = gender_colors) +
  scale_pattern_fill_manual(values = gender_colors) +
  scale_x_discrete(labels = gender_labels) +
  scale_pattern_manual(
    values = c("August" = "none", "December" = "stripe"),
    name = "Month"
  ) +
  guides(
    fill = "none",
    pattern_fill = "none",
    pattern = guide_legend(
      override.aes = list(
        fill = "grey70",
        pattern_density = 0.5,
        pattern_spacing = 0.02
      )
    )
  ) +
  labs(
    x = NULL,
    y = "Resting Heart Rate (bpm)"
  ) +
  theme_seasonality()

p_s3b

# ---- Linear Mixed Model: Sex × Month ----
model_aug_dec_rhr_gender <- lmer(
  RESTING_HEART_RATE ~ GENDER * MONTH_LABEL + splines::ns(AGE_2023, df = 3) + BMI + (1 | USER_ID),
  data = df_aug_dec
)

anova(model_aug_dec_rhr_gender)

# ---- Pairwise comparisons: sex differences within each month ----
emmeans(model_aug_dec_rhr_gender, pairwise ~ GENDER | MONTH_LABEL, adjust = "holm")

# ---- Pairwise comparisons: month differences within each sex ----
emmeans(model_aug_dec_rhr_gender, pairwise ~ MONTH_LABEL | GENDER, adjust = "holm")

# Add within-sex Aug→Dec brackets + interaction-p subtitle
p_s3b <- p_s3b + aug_dec_brackets_layer(
  model_aug_dec_rhr_gender, df_aug_dec, y_var = "RESTING_HEART_RATE",
  by_var = "GENDER", group_levels = c("female", "male"),
  interaction_term = "GENDER:MONTH_LABEL", by_label = "Sex"
)

# Refit model with only main effects
model_rhr_gender_main <- lmer(
  RESTING_HEART_RATE ~ GENDER + MONTH_LABEL + splines::ns(AGE_2023, df = 3) + BMI + (1 | USER_ID),
  data = df_aug_dec
)
anova(model_rhr_gender_main)

# EMMs for gender
emmeans(model_rhr_gender_main, ~ GENDER)

# EMMs for month
emmeans(model_rhr_gender_main, ~ MONTH_LABEL)
```


## HRV by SEX
### Model 1
```{r}
model_hrv_gender <- bam(
  HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = GENDER) + GENDER + s(AGE_2023) + BMI + HRV_person_mean,
  data = df_month,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

v_df_gender_hrv <- predict_marginal_trajectory(
  model_hrv_gender,
  data = df_month,
  by_var = "GENDER"
) %>%
  mutate(GENDER = factor(GENDER, levels = c("male", "female")))

p_fig2c <- ggplot(v_df_gender_hrv, aes(x = MONTH_NUM, y = visregFit, color = GENDER, fill = GENDER)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "Sex", values = gender_colors, labels = gender_labels) +
  scale_fill_manual(name = "Sex", values = gender_colors, labels = gender_labels) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "HRV Offset (ms)") +
  theme_seasonality()

p_fig2c

formula_hrv_sex <- HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = GENDER) + GENDER + s(AGE_2023) + BMI + HRV_person_mean

boot_hrv_sex <- bootstrap_amplitude(df_month, formula_hrv_sex, by = "GENDER")

bootstrap_summary_sex_hrv <- boot_hrv_sex %>%
  group_by(GENDER) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

boot_pvals_sex_hrv <- bootstrap_pairwise(boot_hrv_sex, "GENDER")
```
### Model 2
```{r}
model_hrv_gender_mixed <- bam(
  HRV_within ~ 
    s(MONTH_NUM, bs = "cc", k = 12, by = GENDER) +
    GENDER + s(AGE_2023) + BMI + HRV_person_mean +
    s(USER_ID, bs = "re"),
  data = df_sub,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

v_mixed_gender_df_hrv <- predict_marginal_trajectory(
  model_hrv_gender_mixed,
  data = df_sub,
  by_var = "GENDER",
  exclude_re = "s(USER_ID)"
) %>%
  mutate(GENDER = factor(GENDER, levels = c("male", "female")))

p_s3c <- ggplot(v_mixed_gender_df_hrv, aes(x = MONTH_NUM, y = visregFit, color = GENDER, fill = GENDER)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "Sex", values = gender_colors, labels = gender_labels) +
  scale_fill_manual(name = "Sex", values = gender_colors, labels = gender_labels) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "HRV Offset (ms)") +
  theme_seasonality()

p_s3c

amplitude_gamm_gender_hrv <- v_mixed_gender_df_hrv %>%
  group_by(GENDER) %>%
  summarise(
    amplitude = max(visregFit, na.rm = TRUE) - min(visregFit, na.rm = TRUE),
    peak_month = MONTH_NUM[which.max(visregFit)],
    trough_month = MONTH_NUM[which.min(visregFit)],
    .groups = "drop"
  )

# --- HRV-by-Gender EMMEANS comparisons via lpmatrix ---
res_gender_hrv <- emmeans_contrasts_lpmatrix(model_hrv_gender_mixed, data = df_sub, by_var = "GENDER")
emm_manual_gender_hrv <- res_gender_hrv$emm
gender_contrasts_hrv  <- res_gender_hrv$contrasts

print("Estimated Marginal Means (HRV by Gender):")
print(emm_manual_gender_hrv)
print("Gender Contrasts (HRV):")
print(gender_contrasts_hrv)
```
### Model 3
```{r}
df_amplitude_gender_hrv <- df_month %>%
  filter(MONTH_NUM %in% c(8, 12)) %>%
  mutate(SEASON = case_when(
    MONTH_NUM == 8 ~ "August",
    MONTH_NUM == 12 ~ "December"
  )) %>%
  group_by(USER_ID, SEASON) %>%
  summarise(
    mean_hrv = mean(HRV_within, na.rm = TRUE),
    GENDER = first(GENDER),
    AGE_2023 = first(AGE_2023),
    BMI = first(BMI),
    HRV_person_mean = first(HRV_person_mean),
    .groups = "drop"
  ) %>%
  pivot_wider(names_from = SEASON, values_from = mean_hrv) %>%
  mutate(
    amplitude = August - December
  )

amplitude_summary_gender_hrv <- df_amplitude_gender_hrv %>%
  group_by(GENDER) %>%
  summarise(
    mean_amplitude = mean(amplitude, na.rm = TRUE),
    sd_amplitude = sd(amplitude, na.rm = TRUE),
    n = n(),
    mean_sd = paste0(round(mean_amplitude, 2), " ± ", round(sd_amplitude, 2)),
    .groups = "drop"
  )

model_amplitude_gender_hrv <- lm(amplitude ~ GENDER + splines::ns(AGE_2023, df = 3) + BMI + HRV_person_mean, data = df_amplitude_gender_hrv)
anova(model_amplitude_gender_hrv)
confint(model_amplitude_gender_hrv)

emm_gender_hrv <- emmeans(model_amplitude_gender_hrv, ~ GENDER)
tukey_gender_hrv <- pairs(emm_gender_hrv, adjust = "tukey")
summary(tukey_gender_hrv, infer = c(TRUE, TRUE))

df_amplitude_gender_hrv <- df_amplitude_gender_hrv %>%
  mutate(GENDER = factor(GENDER, levels = c("male", "female")))

p_fig2d <- ggplot(df_amplitude_gender_hrv, aes(x = GENDER, y = amplitude, fill = GENDER)) +
  geom_violin(trim = FALSE, alpha = 0.5, color = NA) +
  geom_boxplot(width = 0.1, outlier.shape = NA, alpha = 0.7) +
  stat_summary(fun = mean, geom = "point", shape = 21, size = 3, fill = "white") +
  scale_fill_manual(values = gender_colors) +
  scale_x_discrete(labels = gender_labels) +
  scale_y_continuous(name = "HRV Amplitude (Aug - Dec, ms)") +
  coord_cartesian(ylim = c(-20, 35)) +
  labs(x = NULL, y = "HRV Amplitude (Aug - Dec, ms)") +
  theme_seasonality(legend = FALSE) +
  violin_brackets_layer(tukey_gender_hrv,
    comparisons = list(c("male", "female")),
    group_levels = c("male", "female"),
    y_positions = c(31),
    tip_length = 1)

p_fig2d
```
### Model 4 - Box and Whisker of Raw HRV Values in Peak and Trough Months by Sex
```{r}
# ---- Filter and prepare data ----
df_aug_dec_hrv <- df_month %>%
  filter(MONTH_NUM %in% c(8, 12)) %>%
  mutate(
    MONTH_LABEL = factor(ifelse(MONTH_NUM == 8, "August", "December"),
                         levels = c("August", "December")),
    GENDER = factor(GENDER, levels = c("male", "female")),
    pattern = ifelse(MONTH_LABEL == "December", "stripe", "none")
  )

# ---- Boxplot: HRV by Sex and Month ----
p_s3d <- ggplot(df_aug_dec_hrv, aes(x = GENDER, y = HRV)) +
  geom_boxplot_pattern(
    aes(
      fill = GENDER,
      pattern = MONTH_LABEL,
      pattern_fill = GENDER
    ),
    position = position_dodge(width = 0.75),
    color = "black",
    pattern_density = 0.4,
    pattern_spacing = 0.05,
    alpha = 0.7,
    outlier.shape = NA
  ) +
  stat_summary(
    fun = median,
    geom = "point",
    aes(group = interaction(MONTH_LABEL, GENDER)),
    position = position_dodge(width = 0.75),
    shape = 21,
    size = 3,
    fill = "white",
    color = "black"
  ) +
  scale_fill_manual(values = gender_colors) +
  scale_pattern_fill_manual(values = gender_colors) +
  scale_x_discrete(labels = gender_labels) +
  scale_pattern_manual(
    values = c("August" = "none", "December" = "stripe"),
    name = "Month"
  ) +
  guides(
    fill = "none",
    pattern_fill = "none",
    pattern = guide_legend(
      override.aes = list(
        fill = "grey70",
        pattern_density = 0.5,
        pattern_spacing = 0.02
      )
    )
  ) +
  labs(
    x = NULL,
    y = "Heart Rate Variability (ms)"
  ) +
  coord_cartesian(ylim = c(0, 150)) +
  theme_seasonality()

p_s3d

# ---- Full model with interaction ----
model_hrv_gender_interact <- lmer(
  HRV ~ GENDER * MONTH_LABEL + splines::ns(AGE_2023, df = 3) + BMI + (1 | USER_ID),
  data = df_aug_dec_hrv
)

anova(model_hrv_gender_interact)

# ---- Pairwise comparisons: GENDER within MONTH ----
emmeans(model_hrv_gender_interact, pairwise ~ GENDER | MONTH_LABEL, adjust = "holm")

# ---- Pairwise comparisons: MONTH within GENDER ----
emmeans(model_hrv_gender_interact, pairwise ~ MONTH_LABEL | GENDER, adjust = "holm")

# Add within-sex Aug→Dec brackets + interaction-p subtitle
p_s3d <- p_s3d + aug_dec_brackets_layer(
  model_hrv_gender_interact, df_aug_dec_hrv, y_var = "HRV",
  by_var = "GENDER", group_levels = c("female", "male"),
  interaction_term = "GENDER:MONTH_LABEL", by_label = "Sex"
)

# ---- Refit model without interaction ----
model_hrv_gender_main <- lmer(
  HRV ~ GENDER + MONTH_LABEL + splines::ns(AGE_2023, df = 3) + BMI + (1 | USER_ID),
  data = df_aug_dec_hrv
)

# ---- Main effect significance ----
anova(model_hrv_gender_main)

# ---- Estimated marginal means ----
emmeans(model_hrv_gender_main, ~ GENDER)
emmeans(model_hrv_gender_main, ~ MONTH_LABEL)

```

## Sex Sensitivity: Age-Matched Subsample (Young + Middle only)
```{r}
# Motivation: female subsample is ~4 yrs younger on average and contains
# roughly half as many older adults as the male subsample. Although all
# primary models adjust for age, restrict here to the age-matched
# young + middle strata (3,000 per age x sex cell) to test whether the
# population-level sex signal survives a cleanly balanced comparison.

df_month_ym <- df_month %>% filter(AGE_BUCKET %in% c("Young", "Middle"))

# Confirm balance
df_month_ym %>% distinct(USER_ID, GENDER, AGE_BUCKET) %>% count(GENDER, AGE_BUCKET)

# ---- RHR: age-matched sex GAM + bootstrap ----
formula_rhr_sex_ym <- RHR_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = GENDER) +
  GENDER + AGE_2023 + BMI + RHR_person_mean

model_rhr_sex_ym <- bam(formula_rhr_sex_ym, data = df_month_ym,
                        method = "fREML", knots = MONTH_KNOTS, discrete = TRUE)

boot_rhr_sex_ym <- bootstrap_amplitude(df_month_ym, formula_rhr_sex_ym, by = "GENDER")

bootstrap_summary_rhr_sex_ym <- boot_rhr_sex_ym %>%
  group_by(GENDER) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low   = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high  = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups  = "drop"
  )

boot_pvals_rhr_sex_ym <- bootstrap_pairwise(boot_rhr_sex_ym, "GENDER")

# ---- HRV: age-matched sex GAM + bootstrap ----
formula_hrv_sex_ym <- HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = GENDER) +
  GENDER + AGE_2023 + BMI + HRV_person_mean

model_hrv_sex_ym <- bam(formula_hrv_sex_ym, data = df_month_ym,
                        method = "fREML", knots = MONTH_KNOTS, discrete = TRUE)

boot_hrv_sex_ym <- bootstrap_amplitude(df_month_ym, formula_hrv_sex_ym, by = "GENDER")

bootstrap_summary_hrv_sex_ym <- boot_hrv_sex_ym %>%
  group_by(GENDER) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low   = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high  = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups  = "drop"
  )

boot_pvals_hrv_sex_ym <- bootstrap_pairwise(boot_hrv_sex_ym, "GENDER")

# ---- Summary table: full sample vs age-matched subsample ----
sex_sensitivity <- bind_rows(
  bootstrap_summary_sex_rhr    %>% mutate(Outcome = "RHR", Sample = "Full"),
  bootstrap_summary_rhr_sex_ym %>% mutate(Outcome = "RHR", Sample = "Young+Middle"),
  bootstrap_summary_sex_hrv    %>% mutate(Outcome = "HRV", Sample = "Full"),
  bootstrap_summary_hrv_sex_ym %>% mutate(Outcome = "HRV", Sample = "Young+Middle")
)

sex_sensitivity_pvals <- bind_rows(
  boot_pvals_sex_rhr       %>% mutate(Outcome = "RHR", Sample = "Full"),
  boot_pvals_rhr_sex_ym    %>% mutate(Outcome = "RHR", Sample = "Young+Middle"),
  boot_pvals_sex_hrv       %>% mutate(Outcome = "HRV", Sample = "Full"),
  boot_pvals_hrv_sex_ym    %>% mutate(Outcome = "HRV", Sample = "Young+Middle")
)

print(sex_sensitivity)
print(sex_sensitivity_pvals)
```


# AIM 2
## Physical Activity
### RHR
```{r}
# Reuse baseline RHR model from Aim 1 (model_rhr_baseline)

# Adjusted model (with PA)
model_rhr_adj_pa <- bam(
  RHR_within ~ s(MONTH_NUM, bs = "cc", k = 12) + s(AGE_2023) + GENDER + BMI + RHR_person_mean + SHRZS_b + SHRZS_w,
  data = df_month,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# ---- Extract smooth fits ----
v_base_df <- predict_marginal_trajectory(model_rhr_baseline, data = df_month) %>% mutate(Model = "Baseline")
v_adj_df  <- predict_marginal_trajectory(model_rhr_adj_pa,  data = df_month) %>% mutate(Model = "Adjusted")

plot_df <- bind_rows(v_base_df, v_adj_df)

# ---- Overlay plot (Figure 3d) ----
p_fig3d <- ggplot(plot_df, aes(x = MONTH_NUM, y = visregFit, color = Model, fill = Model)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(values = c("Baseline" = "black", "Adjusted" = "#1f78b4")) +
  scale_fill_manual(values = c("Baseline" = "black", "Adjusted" = "#1f78b4")) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "RHR Offset (bpm)") +
  theme_seasonality()
p_fig3d

# ---- Bootstrap attenuation ----
att_rhr_pa <- bootstrap_attenuation(
  df_month,
  formula_base = formula_rhr_baseline,
  formula_adj = RHR_within ~ s(MONTH_NUM, bs = "cc", k = 12) + s(AGE_2023) + GENDER + BMI + RHR_person_mean + SHRZS_b + SHRZS_w
)
att_summary_rhr_pa <- summarize_attenuation(att_rhr_pa)
```
### HRV
```{r}
# Reuse baseline HRV model from Aim 1 (model_hrv_baseline)

# Adjusted model (with PA)
model_hrv_adj_pa <- bam(
  HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12) + s(AGE_2023) + GENDER + BMI + HRV_person_mean + SHRZS_b + SHRZS_w,
  data = df_month,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# ---- Extract smooth fits ----
v_base_df_hrv <- predict_marginal_trajectory(model_hrv_baseline, data = df_month) %>% mutate(Model = "Baseline")
v_adj_df_hrv  <- predict_marginal_trajectory(model_hrv_adj_pa,  data = df_month) %>% mutate(Model = "Adjusted")

plot_df_hrv <- bind_rows(v_base_df_hrv, v_adj_df_hrv)

# ---- Overlay plot (Figure 3g) ----
p_fig3g <- ggplot(plot_df_hrv, aes(x = MONTH_NUM, y = visregFit, color = Model, fill = Model)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(values = c("Baseline" = "black", "Adjusted" = "#1f78b4")) +
  scale_fill_manual(values = c("Baseline" = "black", "Adjusted" = "#1f78b4")) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "HRV Offset (ms)") +
  theme_seasonality()
p_fig3g

# ---- Bootstrap attenuation ----
att_hrv_pa <- bootstrap_attenuation(
  df_month,
  formula_base = formula_hrv_baseline,
  formula_adj = HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12) + s(AGE_2023) + GENDER + BMI + HRV_person_mean + SHRZS_b + SHRZS_w
)
att_summary_hrv_pa <- summarize_attenuation(att_hrv_pa)
```

## Sleep
### RHR
```{r}
# Reuse baseline RHR model from Aim 1 (model_rhr_baseline)

# Adjusted model (with sleep)
model_rhr_adj_sleep <- bam(
  RHR_within ~ s(MONTH_NUM, bs = "cc", k = 12) +
    s(AGE_2023) + GENDER + BMI + RHR_person_mean +
    MINUTES_OF_SLEEP_b + MINUTES_OF_SLEEP_w,
  data = df_month,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# ---- Extract smooth fits ----
v_base_df <- predict_marginal_trajectory(model_rhr_baseline, data = df_month) %>% mutate(Model = "Baseline")
v_adj_df  <- predict_marginal_trajectory(model_rhr_adj_sleep, data = df_month) %>% mutate(Model = "Adjusted")

plot_df <- bind_rows(v_base_df, v_adj_df)

# ---- Overlay plot (Figure 3e) ----
p_fig3e <- ggplot(plot_df, aes(x = MONTH_NUM, y = visregFit, color = Model, fill = Model)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(values = c("Baseline" = "black", "Adjusted" = "darkorange")) +
  scale_fill_manual(values = c("Baseline" = "black", "Adjusted" = "darkorange")) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "RHR Offset (bpm)") +
  theme_seasonality()
p_fig3e

# ---- Bootstrap attenuation ----
att_rhr_sleep <- bootstrap_attenuation(
  df_month,
  formula_base = formula_rhr_baseline,
  formula_adj = RHR_within ~ s(MONTH_NUM, bs = "cc", k = 12) + s(AGE_2023) + GENDER + BMI + RHR_person_mean + MINUTES_OF_SLEEP_b + MINUTES_OF_SLEEP_w
)
att_summary_rhr_sleep <- summarize_attenuation(att_rhr_sleep)
```
### HRV
```{r}
# Reuse baseline HRV model from Aim 1 (model_hrv_baseline)

# Adjusted model (with sleep)
model_hrv_adj_sleep <- bam(
  HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12) +
    s(AGE_2023) + GENDER + BMI + HRV_person_mean +
    MINUTES_OF_SLEEP_b + MINUTES_OF_SLEEP_w,
  data = df_month,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# ---- Extract smooth fits ----
v_base_df <- predict_marginal_trajectory(model_hrv_baseline, data = df_month) %>% mutate(Model = "Baseline")
v_adj_df  <- predict_marginal_trajectory(model_hrv_adj_sleep, data = df_month) %>% mutate(Model = "Adjusted")

plot_df <- bind_rows(v_base_df, v_adj_df)

# ---- Overlay plot (Figure 3h) ----
p_fig3h <- ggplot(plot_df, aes(x = MONTH_NUM, y = visregFit, color = Model, fill = Model)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(values = c("Baseline" = "black", "Adjusted" = "darkorange")) +
  scale_fill_manual(values = c("Baseline" = "black", "Adjusted" = "darkorange")) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "HRV Offset (ms)") +
  theme_seasonality()
p_fig3h

# ---- Bootstrap attenuation ----
att_hrv_sleep <- bootstrap_attenuation(
  df_month,
  formula_base = formula_hrv_baseline,
  formula_adj = HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12) + s(AGE_2023) + GENDER + BMI + HRV_person_mean + MINUTES_OF_SLEEP_b + MINUTES_OF_SLEEP_w
)
att_summary_hrv_sleep <- summarize_attenuation(att_hrv_sleep)
```

# AIM 3
### RHR
```{r}
# --- 1. Fit adjusted GAMM ---
model_rhr_age_alladj <- bam(
  RHR_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = AGE_BUCKET) + AGE_BUCKET +
    GENDER + BMI + RHR_person_mean +
    SHRZS_b + SHRZS_w +
    MINUTES_OF_SLEEP_b + MINUTES_OF_SLEEP_w,
  data = df_month,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# --- 2. Extract smooths via marginal-trajectory helper ---
v_rhr_df <- predict_marginal_trajectory(
  model_rhr_age_alladj,
  data = df_month,
  by_var = "AGE_BUCKET"
) %>%
  mutate(AGE_BUCKET = factor(AGE_BUCKET, levels = c("Young", "Middle", "Old")))

# Figure S5 panel A
p_s5a <- ggplot(v_rhr_df, aes(x = MONTH_NUM, y = visregFit, color = AGE_BUCKET, fill = AGE_BUCKET)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_color_manual(name = "Age Group", values = age_colors) +
  scale_fill_manual(name = "Age Group", values = age_colors) +
  scale_x_continuous(
    breaks = 1:12,
    labels = month.abb,
    name = NULL
  ) +
  scale_y_continuous(name = "RHR Offset (bpm)") +
  theme_seasonality()
p_s5a

# ---- Bootstrap modeled amplitude by AGE_BUCKET (adjusted) ----
formula_rhr_age_adj <- RHR_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = AGE_BUCKET) + AGE_BUCKET +
  GENDER + BMI + RHR_person_mean +
  SHRZS_b + SHRZS_w +
  MINUTES_OF_SLEEP_b + MINUTES_OF_SLEEP_w

boot_rhr_age_adj <- bootstrap_amplitude(df_month, formula_rhr_age_adj, by = "AGE_BUCKET")

bootstrap_summary_rhr <- boot_rhr_age_adj %>%
  group_by(AGE_BUCKET) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

boot_pvals_rhr <- bootstrap_pairwise(boot_rhr_age_adj, "AGE_BUCKET")
```
### HRV
```{r}
model_hrv_age_alladj <- bam(
  HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = AGE_BUCKET) + AGE_BUCKET +
    GENDER + BMI + HRV_person_mean +
    SHRZS_b + SHRZS_w +
    MINUTES_OF_SLEEP_b + MINUTES_OF_SLEEP_w,
  data = df_month,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

v_hrv_df <- predict_marginal_trajectory(
  model_hrv_age_alladj,
  data = df_month,
  by_var = "AGE_BUCKET"
) %>%
  mutate(AGE_BUCKET = factor(AGE_BUCKET, levels = c("Young", "Middle", "Old")))

# Figure S5 panel B
p_s5b <- ggplot(v_hrv_df, aes(x = MONTH_NUM, y = visregFit, color = AGE_BUCKET, fill = AGE_BUCKET)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_color_manual(name = "Age Group", values = age_colors) +
  scale_fill_manual(name = "Age Group", values = age_colors) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "HRV Offset (ms)") +
  theme_seasonality()
p_s5b

# ---- Bootstrap modeled amplitude by AGE_BUCKET (adjusted) ----
formula_hrv_age_adj <- HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = AGE_BUCKET) + AGE_BUCKET +
  GENDER + BMI + HRV_person_mean +
  SHRZS_b + SHRZS_w +
  MINUTES_OF_SLEEP_b + MINUTES_OF_SLEEP_w

boot_hrv_age_adj <- bootstrap_amplitude(df_month, formula_hrv_age_adj, by = "AGE_BUCKET")

bootstrap_summary_hrv_adj <- boot_hrv_age_adj %>%
  group_by(AGE_BUCKET) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

boot_pvals_hrv_adj <- bootstrap_pairwise(boot_hrv_age_adj, "AGE_BUCKET")
```


# AIM 4
### AIM 4: Data Preparation (PA Tertiles + BMI Categories)
```{r}
# ---- 1. Add PA Tertiles to df_month (once; used by all PA analyses) ----
# Tertile is computed at the user level (one row per user) and joined back, so
# users whose SHRZS_b ties at a cutpoint are not split across two tertiles.
df_month <- df_month %>%
  select(-any_of("PA_TERTILE")) %>%
  mutate(
    AGE_BUCKET = as.factor(AGE_BUCKET),
    GENDER = as.factor(GENDER)
  )

user_pa <- df_month %>%
  distinct(USER_ID, SHRZS_b) %>%
  mutate(PA_TERTILE = factor(ntile(SHRZS_b, 3),
                             labels = c("Low", "Moderate", "High"))) %>%
  select(USER_ID, PA_TERTILE)  # drop SHRZS_b before join to avoid .x/.y collision

df_month <- df_month %>%
  select(-any_of(c("SHRZS_b.x", "SHRZS_b.y", "PA_TERTILE.x", "PA_TERTILE.y"))) %>%
  left_join(user_pa, by = "USER_ID")

# Verification: cell counts should total 16,415 (one row per user)
df_month %>%
  distinct(USER_ID, AGE_BUCKET, PA_TERTILE) %>%
  count(AGE_BUCKET, PA_TERTILE)

# ---- 2. Create df_month_bmi (separate copy; preserves df_month intact) ----
# Underweight (BMI < 18.5) and missing-BMI participants receive NA from
# case_when() and are dropped via drop_na() below.
df_month_bmi <- df_month %>%
  mutate(
    BMI_CATEGORY = case_when(
      BMI >= 18.5 & BMI < 25 ~ "Normal",
      BMI >= 25 & BMI < 30 ~ "Overweight",
      BMI >= 30 ~ "Obese"
    ),
    BMI_CATEGORY = factor(BMI_CATEGORY, levels = c("Normal", "Overweight", "Obese"))
  ) %>%
  drop_na(BMI_CATEGORY)
```

### RHR Moderation by Activity: All
```{r}
# ---- 1. Fit GAMM with cyclic spline by PA_TERTILE ----
formula_rhr_pa_all <- RHR_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = PA_TERTILE) +
  PA_TERTILE + s(AGE_2023) + GENDER + BMI + RHR_person_mean

model_rhr_pa_all <- bam(
  formula_rhr_pa_all,
  data = df_month,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# ---- 2. Marginal trajectory (averages over GENDER; lpmatrix-based SE) ----
v_df_pa_all <- predict_marginal_trajectory(
  model_rhr_pa_all,
  data = df_month,
  by_var = "PA_TERTILE"
)

# ---- 4. Plot ----
p_fig4a <- ggplot(v_df_pa_all, aes(x = MONTH_NUM, y = visregFit, color = PA_TERTILE, fill = PA_TERTILE)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "Activity Group", values = pa_colors) +
  scale_fill_manual(name = "Activity Group", values = pa_colors) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "RHR Offset (bpm)") +
  theme_seasonality()
p_fig4a

# ---- 5. Bootstrap amplitude + pairwise ----
boot_rhr_pa_all <- bootstrap_amplitude(df_month, formula_rhr_pa_all, by = "PA_TERTILE")

bootstrap_summary_pa_all <- boot_rhr_pa_all %>%
  group_by(PA_TERTILE) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

boot_pvals_pa_all <- bootstrap_pairwise(boot_rhr_pa_all, "PA_TERTILE")
```

### HRV Moderation by Activity: All
```{r}
# ---- 1. Fit GAMM with cyclic spline by PA_TERTILE ----
formula_hrv_pa_all <- HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = PA_TERTILE) +
  PA_TERTILE + s(AGE_2023) + GENDER + BMI + HRV_person_mean

model_hrv_pa_all <- bam(
  formula_hrv_pa_all,
  data = df_month,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# ---- 2. Marginal trajectory (averages over GENDER; lpmatrix-based SE) ----
v_df_hrv_pa_all <- predict_marginal_trajectory(
  model_hrv_pa_all,
  data = df_month,
  by_var = "PA_TERTILE"
)

# ---- 4. Plot ----
p_fig4b <- ggplot(v_df_hrv_pa_all, aes(x = MONTH_NUM, y = visregFit, color = PA_TERTILE, fill = PA_TERTILE)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "Activity Group", values = pa_colors) +
  scale_fill_manual(name = "Activity Group", values = pa_colors) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "HRV Offset (ms)") +
  theme_seasonality()
p_fig4b

# ---- 5. Bootstrap amplitude + pairwise ----
boot_hrv_pa_all <- bootstrap_amplitude(df_month, formula_hrv_pa_all, by = "PA_TERTILE")

bootstrap_summary_hrv_pa_all <- boot_hrv_pa_all %>%
  group_by(PA_TERTILE) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

boot_pvals_hrv_pa_all <- bootstrap_pairwise(boot_hrv_pa_all, "PA_TERTILE")
```

### RHR Moderation by BMI: All
```{r}
# ---- 1. Fit GAMM with cyclic spline by BMI_CATEGORY (uses df_month_bmi) ----
formula_rhr_bmi_all <- RHR_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = BMI_CATEGORY) +
  BMI_CATEGORY + s(AGE_2023) + SHRZS_b + SHRZS_w + GENDER + RHR_person_mean

model_rhr_bmi_all <- bam(
  formula_rhr_bmi_all,
  data = df_month_bmi,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# ---- 2. Marginal trajectory (averages over GENDER; lpmatrix-based SE) ----
v_df_bmi_all <- predict_marginal_trajectory(
  model_rhr_bmi_all,
  data = df_month_bmi,
  by_var = "BMI_CATEGORY"
)

# ---- 4. Plot ----
p_fig4c <- ggplot(v_df_bmi_all, aes(x = MONTH_NUM, y = visregFit, color = BMI_CATEGORY, fill = BMI_CATEGORY)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "BMI Group", values = bmi_colors) +
  scale_fill_manual(name = "BMI Group", values = bmi_colors) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "RHR Offset (bpm)") +
  theme_seasonality()
p_fig4c

# ---- 5. Bootstrap amplitude + pairwise (formula matches primary model) ----
boot_rhr_bmi_all <- bootstrap_amplitude(df_month_bmi, formula_rhr_bmi_all, by = "BMI_CATEGORY")

bootstrap_summary_bmi_all <- boot_rhr_bmi_all %>%
  group_by(BMI_CATEGORY) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

boot_pvals_bmi_all <- bootstrap_pairwise(boot_rhr_bmi_all, "BMI_CATEGORY")
```

### HRV Moderation by BMI: All
```{r}
# ---- 1. Fit GAMM with cyclic spline by BMI_CATEGORY (uses df_month_bmi) ----
formula_hrv_bmi_all <- HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = BMI_CATEGORY) +
  BMI_CATEGORY + s(AGE_2023) + SHRZS_b + SHRZS_w + GENDER + HRV_person_mean

model_hrv_bmi_all <- bam(
  formula_hrv_bmi_all,
  data = df_month_bmi,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# ---- 2. Marginal trajectory (averages over GENDER; lpmatrix-based SE) ----
v_df_hrv_bmi_all <- predict_marginal_trajectory(
  model_hrv_bmi_all,
  data = df_month_bmi,
  by_var = "BMI_CATEGORY"
)

# ---- 4. Plot ----
p_fig4d <- ggplot(v_df_hrv_bmi_all, aes(x = MONTH_NUM, y = visregFit, color = BMI_CATEGORY, fill = BMI_CATEGORY)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "BMI Group", values = bmi_colors) +
  scale_fill_manual(name = "BMI Group", values = bmi_colors) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "HRV Offset (ms)") +
  theme_seasonality()
p_fig4d

# ---- 5. Bootstrap amplitude + pairwise (formula matches primary model) ----
boot_hrv_bmi_all <- bootstrap_amplitude(df_month_bmi, formula_hrv_bmi_all, by = "BMI_CATEGORY")

bootstrap_summary_hrv_bmi_all <- boot_hrv_bmi_all %>%
  group_by(BMI_CATEGORY) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

boot_pvals_hrv_bmi_all <- bootstrap_pairwise(boot_hrv_bmi_all, "BMI_CATEGORY")
```


### Geographic Daylight Amplitude: Regression + Figures
```{r}
# State-centroid latitudes (built-in to base R)
state_lat <- tibble::tibble(STATE = state.abb, LAT = state.center$y)

# Solar declination + daylight hours from latitude (degrees N) and day-of-year
solar_declination <- function(d) asin(0.39795 * cos(0.98563 * pi/180 * (d - 173)))
daylight_hours <- function(lat_deg, day_of_year) {
  phi   <- lat_deg * pi/180
  delta <- solar_declination(day_of_year)
  H     <- acos(pmin(pmax((-0.01454 - sin(phi) * sin(delta)) / (cos(phi) * cos(delta)),
                          -1), 1))
  2 * H * 180/pi / 15
}

# Annual daylight amplitude = summer solstice (day 173) − winter solstice (day 355)
state_lat <- state_lat %>%
  mutate(DAYLIGHT_AMP = daylight_hours(LAT, 173) - daylight_hours(LAT, 355))

# ---- User-level Aug/Dec amplitudes joined to daylight amplitude (RHR) ----
df_amp_rhr_lat <- df_month %>%
  filter(MONTH_NUM %in% c(8, 12)) %>%
  mutate(SEASON = ifelse(MONTH_NUM == 8, "August", "December")) %>%
  group_by(USER_ID, SEASON) %>%
  summarise(
    mean_val = mean(RHR_within, na.rm = TRUE),
    STATE = first(na.omit(STATE)),
    AGE_2023 = first(AGE_2023),
    GENDER = first(GENDER),
    BMI = first(BMI),
    RHR_person_mean = first(RHR_person_mean),
    .groups = "drop"
  ) %>%
  pivot_wider(names_from = SEASON, values_from = mean_val) %>%
  mutate(amplitude = December - August) %>%
  left_join(state_lat, by = "STATE") %>%
  drop_na(DAYLIGHT_AMP, amplitude)

# ---- User-level Aug/Dec amplitudes joined to daylight amplitude (HRV) ----
df_amp_hrv_lat <- df_month %>%
  filter(MONTH_NUM %in% c(8, 12)) %>%
  mutate(SEASON = ifelse(MONTH_NUM == 8, "August", "December")) %>%
  group_by(USER_ID, SEASON) %>%
  summarise(
    mean_val = mean(HRV_within, na.rm = TRUE),
    STATE = first(na.omit(STATE)),
    AGE_2023 = first(AGE_2023),
    GENDER = first(GENDER),
    BMI = first(BMI),
    HRV_person_mean = first(HRV_person_mean),
    .groups = "drop"
  ) %>%
  pivot_wider(names_from = SEASON, values_from = mean_val) %>%
  mutate(amplitude = August - December) %>%
  left_join(state_lat, by = "STATE") %>%
  drop_na(DAYLIGHT_AMP, amplitude)

# ---- Linear regressions: amplitude ~ annual daylight amplitude + covariates ----
model_rhr_daylight <- lm(amplitude ~ DAYLIGHT_AMP + AGE_2023 + GENDER + BMI + RHR_person_mean,
                         data = df_amp_rhr_lat)
model_hrv_daylight <- lm(amplitude ~ DAYLIGHT_AMP + AGE_2023 + GENDER + BMI + HRV_person_mean,
                         data = df_amp_hrv_lat)

# Extract daylight-amplitude coefficient, CI, and p-value for reporting
daylight_results <- tibble::tibble(
  Outcome       = c("RHR", "HRV"),
  beta_daylight = c(coef(model_rhr_daylight)["DAYLIGHT_AMP"],
                    coef(model_hrv_daylight)["DAYLIGHT_AMP"]),
  ci_low        = c(confint(model_rhr_daylight)["DAYLIGHT_AMP", 1],
                    confint(model_hrv_daylight)["DAYLIGHT_AMP", 1]),
  ci_high       = c(confint(model_rhr_daylight)["DAYLIGHT_AMP", 2],
                    confint(model_hrv_daylight)["DAYLIGHT_AMP", 2]),
  p_value       = c(summary(model_rhr_daylight)$coefficients["DAYLIGHT_AMP", "Pr(>|t|)"],
                    summary(model_hrv_daylight)$coefficients["DAYLIGHT_AMP", "Pr(>|t|)"]),
  n             = c(nrow(df_amp_rhr_lat), nrow(df_amp_hrv_lat))
)
print(daylight_results)

# ---- Sensitivity: cluster-robust SEs by state ----
# DAYLIGHT_AMP is identical for all users in a state, so OLS treats clustered
# observations as independent. Cluster-robust standard errors (sandwich
# estimator) account for within-state correlation without imposing
# distributional assumptions on state-level effects.
library(sandwich)
library(lmtest)

vcov_cluster_rhr <- vcovCL(model_rhr_daylight, cluster = ~ STATE)
vcov_cluster_hrv <- vcovCL(model_hrv_daylight, cluster = ~ STATE)

coeftest_rhr_cluster <- coeftest(model_rhr_daylight, vcov = vcov_cluster_rhr)
coeftest_hrv_cluster <- coeftest(model_hrv_daylight, vcov = vcov_cluster_hrv)

ci_rhr_cluster <- coefci(model_rhr_daylight, vcov = vcov_cluster_rhr)
ci_hrv_cluster <- coefci(model_hrv_daylight, vcov = vcov_cluster_hrv)

daylight_results_cluster <- tibble::tibble(
  Outcome       = c("RHR", "HRV"),
  beta_daylight = c(coef(model_rhr_daylight)["DAYLIGHT_AMP"],
                    coef(model_hrv_daylight)["DAYLIGHT_AMP"]),
  ci_low        = c(ci_rhr_cluster["DAYLIGHT_AMP", 1],
                    ci_hrv_cluster["DAYLIGHT_AMP", 1]),
  ci_high       = c(ci_rhr_cluster["DAYLIGHT_AMP", 2],
                    ci_hrv_cluster["DAYLIGHT_AMP", 2]),
  p_value       = c(coeftest_rhr_cluster["DAYLIGHT_AMP", "Pr(>|t|)"],
                    coeftest_hrv_cluster["DAYLIGHT_AMP", "Pr(>|t|)"]),
  n_users       = c(nrow(df_amp_rhr_lat), nrow(df_amp_hrv_lat)),
  n_states      = c(length(unique(df_amp_rhr_lat$STATE)), length(unique(df_amp_hrv_lat$STATE)))
)
print(daylight_results_cluster)

# Side-by-side comparison: OLS vs cluster-robust SEs (same coefficients, different SEs)
daylight_comparison <- bind_rows(
  daylight_results            %>% select(Outcome, beta_daylight, ci_low, ci_high, p_value) %>% mutate(SE_method = "OLS"),
  daylight_results_cluster    %>% select(Outcome, beta_daylight, ci_low, ci_high, p_value) %>% mutate(SE_method = "Cluster-robust (state)")
) %>%
  select(SE_method, Outcome, beta_daylight, ci_low, ci_high, p_value)
print(daylight_comparison)

# ---- State-level summaries for plotting (one point per state, sized by n) ----
state_summary_rhr <- df_amp_rhr_lat %>%
  group_by(STATE, DAYLIGHT_AMP) %>%
  summarise(mean_amp = mean(amplitude, na.rm = TRUE),
            n_users = dplyr::n(), .groups = "drop")

state_summary_hrv <- df_amp_hrv_lat %>%
  group_by(STATE, DAYLIGHT_AMP) %>%
  summarise(mean_amp = mean(amplitude, na.rm = TRUE),
            n_users = dplyr::n(), .groups = "drop")

# ---- Cluster-robust prediction grids for plotting ----
# Build prediction grids with DAYLIGHT_AMP varying and other covariates at
# sample means (GENDER at reference level "male"). Cluster-robust pointwise
# SEs computed via X %*% vcovCL %*% X'.

# RHR
pred_rhr <- data.frame(
  DAYLIGHT_AMP    = seq(min(df_amp_rhr_lat$DAYLIGHT_AMP),
                        max(df_amp_rhr_lat$DAYLIGHT_AMP), length.out = 200),
  AGE_2023        = mean(df_amp_rhr_lat$AGE_2023, na.rm = TRUE),
  GENDER          = factor(levels(df_amp_rhr_lat$GENDER)[1],
                           levels = levels(df_amp_rhr_lat$GENDER)),
  BMI             = mean(df_amp_rhr_lat$BMI, na.rm = TRUE),
  RHR_person_mean = mean(df_amp_rhr_lat$RHR_person_mean, na.rm = TRUE)
)
X_rhr <- model.matrix(~ DAYLIGHT_AMP + AGE_2023 + GENDER + BMI + RHR_person_mean,
                      data = pred_rhr)
pred_rhr$fit <- as.vector(X_rhr %*% coef(model_rhr_daylight))
pred_rhr$se  <- sqrt(diag(X_rhr %*% vcov_cluster_rhr %*% t(X_rhr)))
pred_rhr$lwr <- pred_rhr$fit - 1.96 * pred_rhr$se
pred_rhr$upr <- pred_rhr$fit + 1.96 * pred_rhr$se

# HRV
pred_hrv <- data.frame(
  DAYLIGHT_AMP    = seq(min(df_amp_hrv_lat$DAYLIGHT_AMP),
                        max(df_amp_hrv_lat$DAYLIGHT_AMP), length.out = 200),
  AGE_2023        = mean(df_amp_hrv_lat$AGE_2023, na.rm = TRUE),
  GENDER          = factor(levels(df_amp_hrv_lat$GENDER)[1],
                           levels = levels(df_amp_hrv_lat$GENDER)),
  BMI             = mean(df_amp_hrv_lat$BMI, na.rm = TRUE),
  HRV_person_mean = mean(df_amp_hrv_lat$HRV_person_mean, na.rm = TRUE)
)
X_hrv <- model.matrix(~ DAYLIGHT_AMP + AGE_2023 + GENDER + BMI + HRV_person_mean,
                      data = pred_hrv)
pred_hrv$fit <- as.vector(X_hrv %*% coef(model_hrv_daylight))
pred_hrv$se  <- sqrt(diag(X_hrv %*% vcov_cluster_hrv %*% t(X_hrv)))
pred_hrv$lwr <- pred_hrv$fit - 1.96 * pred_hrv$se
pred_hrv$upr <- pred_hrv$fit + 1.96 * pred_hrv$se

# ---- Figure S5A (saved as S5; internal name p_s9a per creation order) ----
# Caption note: points are state-level means (sized by n users); regression
# line and 95% CI are from a user-level model with state-clustered SEs.
p_s9a <- ggplot(state_summary_rhr, aes(x = DAYLIGHT_AMP, y = mean_amp)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  geom_ribbon(data = pred_rhr, aes(x = DAYLIGHT_AMP, ymin = lwr, ymax = upr),
              fill = "#D62728", alpha = 0.25, inherit.aes = FALSE) +
  geom_line(data = pred_rhr, aes(x = DAYLIGHT_AMP, y = fit),
            color = "#D62728", linewidth = 1, inherit.aes = FALSE) +
  geom_point(aes(size = n_users), alpha = 0.75, color = "gray25") +
  scale_size_continuous(name = "N per State", range = c(1.5, 8)) +
  scale_x_continuous(name = "Annual Daylight Amplitude (h)") +
  scale_y_continuous(name = "RHR Amplitude (Dec − Aug, bpm)") +
  theme_seasonality()
p_s9a

# ---- Figure S5B (saved as S5; internal name p_s9b per creation order) ----
p_s9b <- ggplot(state_summary_hrv, aes(x = DAYLIGHT_AMP, y = mean_amp)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  geom_ribbon(data = pred_hrv, aes(x = DAYLIGHT_AMP, ymin = lwr, ymax = upr),
              fill = "#1F77B4", alpha = 0.25, inherit.aes = FALSE) +
  geom_line(data = pred_hrv, aes(x = DAYLIGHT_AMP, y = fit),
            color = "#1F77B4", linewidth = 1, inherit.aes = FALSE) +
  geom_point(aes(size = n_users), alpha = 0.75, color = "gray25") +
  scale_size_continuous(name = "N per State", range = c(1.5, 8)) +
  scale_x_continuous(name = "Annual Daylight Amplitude (h)") +
  scale_y_continuous(name = "HRV Amplitude (Aug − Dec, ms)") +
  theme_seasonality()
p_s9b
```


### Old Adults: Create Subsets
```{r}
# ---- Create df_old ONCE (PA analyses) ----
df_old <- df_month %>%
  filter(AGE_BUCKET == "Old")

# ---- Create df_old_bmi ONCE (BMI analyses) ----
df_old_bmi <- df_month_bmi %>%
  filter(AGE_BUCKET == "Old")
```

### RHR Moderation by Activity: Old
```{r}
# ---- 1. Fit GAMM with cyclic spline by PA_TERTILE ----
formula_rhr_pa_old <- RHR_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = PA_TERTILE) +
  PA_TERTILE + GENDER + BMI + RHR_person_mean

model_rhr_pa_old <- bam(
  formula_rhr_pa_old,
  data = df_old,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# ---- 2. Marginal trajectory (averages over GENDER; lpmatrix-based SE) ----
v_df_pa <- predict_marginal_trajectory(
  model_rhr_pa_old,
  data = df_old,
  by_var = "PA_TERTILE"
)

# ---- 5. Plot ----
p_s8_rhr_pa <- ggplot(v_df_pa, aes(x = MONTH_NUM, y = visregFit, color = PA_TERTILE, fill = PA_TERTILE)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "Activity Group", values = pa_colors) +
  scale_fill_manual(name = "Activity Group", values = pa_colors) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "RHR Offset (bpm)") +
  theme_seasonality()
p_s8_rhr_pa

# ---- 6. Bootstrap amplitude + pairwise ----
boot_rhr_pa_old <- bootstrap_amplitude(df_old, formula_rhr_pa_old, by = "PA_TERTILE")

bootstrap_summary_pa_old <- boot_rhr_pa_old %>%
  group_by(PA_TERTILE) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

boot_pvals_pa_old <- bootstrap_pairwise(boot_rhr_pa_old, "PA_TERTILE")

# ---- 7. Plot Bootstrapped Amplitude by PA_TERTILE ----
bootstrap_summary_pa_old$PA_TERTILE <- factor(bootstrap_summary_pa_old$PA_TERTILE, levels = c("Low", "Moderate", "High"))

p_s8_rhr_pa_amp <- ggplot(bootstrap_summary_pa_old, aes(x = PA_TERTILE, y = mean_amp, fill = PA_TERTILE)) +
  geom_col(width = 0.6, color = "black", alpha = 0.8) +
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high), width = 0.2, linewidth = 1.1) +
  scale_fill_manual(values = pa_colors) +
  labs(
    x = "Physical Activity Tertile (Older Adults)",
    y = "Modeled RHR Amplitude (bpm)"
  ) +
  theme_seasonality(legend = FALSE)
p_s8_rhr_pa_amp
```

### HRV Moderation by Activity: Old
```{r}
# ---- 1. Fit GAMM with cyclic spline by PA_TERTILE ----
formula_hrv_pa_old <- HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = PA_TERTILE) +
  PA_TERTILE + GENDER + BMI + HRV_person_mean

model_hrv_pa_old <- bam(
  formula_hrv_pa_old,
  data = df_old,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# ---- 2. Marginal trajectory (averages over GENDER; lpmatrix-based SE) ----
v_df_hrv <- predict_marginal_trajectory(
  model_hrv_pa_old,
  data = df_old,
  by_var = "PA_TERTILE"
)

# ---- 5. Plot ----
p_s8_hrv_pa <- ggplot(v_df_hrv, aes(x = MONTH_NUM, y = visregFit, color = PA_TERTILE, fill = PA_TERTILE)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "Activity Group", values = pa_colors) +
  scale_fill_manual(name = "Activity Group", values = pa_colors) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "HRV Offset (ms)") +
  theme_seasonality()
p_s8_hrv_pa

# ---- 6. Bootstrap amplitude + pairwise ----
boot_hrv_pa_old <- bootstrap_amplitude(df_old, formula_hrv_pa_old, by = "PA_TERTILE")

bootstrap_summary_hrv_pa_old <- boot_hrv_pa_old %>%
  group_by(PA_TERTILE) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

boot_pvals_hrv_pa_old <- bootstrap_pairwise(boot_hrv_pa_old, "PA_TERTILE")
```

### RHR Moderation by BMI: Old
```{r}
df_old_bmi %>%
  group_by(BMI_CATEGORY) %>%
  summarise(n_users = n_distinct(USER_ID), .groups = "drop")

# ---- 2. Fit GAMM with cyclic spline by BMI_CATEGORY ----
formula_rhr_bmi_old <- RHR_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = BMI_CATEGORY) +
  BMI_CATEGORY + GENDER + SHRZS_b + SHRZS_w + RHR_person_mean

model_rhr_bmi_old <- bam(
  formula_rhr_bmi_old,
  data = df_old_bmi,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# ---- 3. Marginal trajectory (averages over GENDER; lpmatrix-based SE) ----
v_df_bmi <- predict_marginal_trajectory(
  model_rhr_bmi_old,
  data = df_old_bmi,
  by_var = "BMI_CATEGORY"
)

# ---- 4. Plot ----
p_s8_rhr_bmi <- ggplot(v_df_bmi, aes(x = MONTH_NUM, y = visregFit, color = BMI_CATEGORY, fill = BMI_CATEGORY)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "BMI Group", values = bmi_colors) +
  scale_fill_manual(name = "BMI Group", values = bmi_colors) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "RHR Offset (bpm)") +
  theme_seasonality()
p_s8_rhr_bmi

# ---- 5. Bootstrap amplitude + pairwise (formula matches primary model) ----
boot_rhr_bmi_old <- bootstrap_amplitude(df_old_bmi, formula_rhr_bmi_old, by = "BMI_CATEGORY")

bootstrap_summary_bmi_old <- boot_rhr_bmi_old %>%
  group_by(BMI_CATEGORY) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

boot_pvals_bmi_old <- bootstrap_pairwise(boot_rhr_bmi_old, "BMI_CATEGORY")
```

### RHR Moderation by BMI: Old (Plotting Only; No Overweight)
```{r}
df_old_bmi_no <- df_old_bmi %>%
  filter(BMI_CATEGORY %in% c("Normal", "Obese")) %>%
  mutate(BMI_CATEGORY = factor(BMI_CATEGORY, levels = c("Normal", "Obese")))

# Fit GAMM with cyclic spline by BMI_CATEGORY
model_rhr_bmi_noow <- bam(
  RHR_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = BMI_CATEGORY) +
    BMI_CATEGORY + GENDER + SHRZS_b + SHRZS_w + BMI + RHR_person_mean,
  data = df_old_bmi_no,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# Marginal trajectory (averages over GENDER; lpmatrix-based SE)
v_df_bmi_noow <- predict_marginal_trajectory(
  model_rhr_bmi_noow,
  data = df_old_bmi_no,
  by_var = "BMI_CATEGORY"
)

# Plot
p_s8_rhr_bmi_noow <- ggplot(v_df_bmi_noow, aes(x = MONTH_NUM, y = visregFit, color = BMI_CATEGORY, fill = BMI_CATEGORY)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "BMI Group", values = bmi_colors[c("Normal", "Obese")]) +
  scale_fill_manual(name = "BMI Group", values = bmi_colors[c("Normal", "Obese")]) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "RHR Offset (bpm)") +
  theme_seasonality()
p_s8_rhr_bmi_noow
```

### HRV Moderation by BMI: Old
```{r}
# ---- 1. Fit GAMM with cyclic spline by BMI_CATEGORY (reuses df_old_bmi) ----
formula_hrv_bmi_old <- HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = BMI_CATEGORY) +
  BMI_CATEGORY + GENDER + SHRZS_b + SHRZS_w + HRV_person_mean

model_hrv_bmi_old <- bam(
  formula_hrv_bmi_old,
  data = df_old_bmi,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# ---- 2. Marginal trajectory (averages over GENDER; lpmatrix-based SE) ----
v_df_hrv_bmi <- predict_marginal_trajectory(
  model_hrv_bmi_old,
  data = df_old_bmi,
  by_var = "BMI_CATEGORY"
)

# ---- 3. Plot ----
p_s8_hrv_bmi <- ggplot(v_df_hrv_bmi, aes(x = MONTH_NUM, y = visregFit, color = BMI_CATEGORY, fill = BMI_CATEGORY)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "BMI Group", values = bmi_colors) +
  scale_fill_manual(name = "BMI Group", values = bmi_colors) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "HRV Offset (ms)") +
  theme_seasonality()
p_s8_hrv_bmi

# ---- 4. Bootstrap amplitude + pairwise (formula matches primary model) ----
boot_hrv_bmi_old <- bootstrap_amplitude(df_old_bmi, formula_hrv_bmi_old, by = "BMI_CATEGORY")

bootstrap_summary_hrv_bmi_old <- boot_hrv_bmi_old %>%
  group_by(BMI_CATEGORY) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

boot_pvals_hrv_bmi_old <- bootstrap_pairwise(boot_hrv_bmi_old, "BMI_CATEGORY")
```

### HRV Moderation by BMI: Old (Plotting Only; No Overweight)
```{r}
# ---- Fit GAMM with cyclic spline by BMI_CATEGORY (reuses df_old_bmi_no from RHR section) ----
model_hrv_bmi_noow <- bam(
  HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = BMI_CATEGORY) +
    BMI_CATEGORY + GENDER + BMI + HRV_person_mean,
  data = df_old_bmi_no,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# ---- 3. Marginal trajectory (averages over GENDER; lpmatrix-based SE) ----
v_df_hrv_bmi_noow <- predict_marginal_trajectory(
  model_hrv_bmi_noow,
  data = df_old_bmi_no,
  by_var = "BMI_CATEGORY"
)

# ---- 4. Plot ----
p_s8_hrv_bmi_noow <- ggplot(v_df_hrv_bmi_noow, aes(x = MONTH_NUM, y = visregFit, color = BMI_CATEGORY, fill = BMI_CATEGORY)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "BMI Group", values = bmi_colors[c("Normal", "Obese")]) +
  scale_fill_manual(name = "BMI Group", values = bmi_colors[c("Normal", "Obese")]) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "HRV Offset (ms)") +
  theme_seasonality()
p_s8_hrv_bmi_noow
```
### RHR Box-And-Whisker and LMM by BMI Category: Old
```{r}
df_old_bmi_aug_dec <- df_old_bmi %>%
  filter(MONTH_NUM %in% c(8, 12)) %>%
  mutate(
    MONTH_LABEL = factor(ifelse(MONTH_NUM == 8, "August", "December"), levels = c("August", "December")),
    pattern = ifelse(MONTH_LABEL == "December", "stripe", "none")
  )

# Boxplot
p_s8_rhr_bmi_box <- ggplot(df_old_bmi_aug_dec, aes(x = BMI_CATEGORY, y = RESTING_HEART_RATE)) +
  geom_boxplot_pattern(
    aes(
      fill = BMI_CATEGORY,
      pattern = MONTH_LABEL,
      pattern_fill = BMI_CATEGORY
    ),
    position = position_dodge(width = 0.75),
    color = "black",
    pattern_density = 0.4,
    pattern_spacing = 0.05,
    alpha = 0.7,
    outlier.shape = NA
  ) +
  stat_summary(
    fun = median,
    geom = "point",
    aes(group = interaction(MONTH_LABEL, BMI_CATEGORY)),
    position = position_dodge(width = 0.75),
    shape = 21,
    size = 3,
    fill = "white",
    color = "black"
  ) +
  scale_fill_manual(values = bmi_colors) +
  scale_pattern_fill_manual(values = bmi_colors) +
  scale_pattern_manual(
    values = c("August" = "none", "December" = "stripe"),
    name = "Month"
  ) +
  guides(
    fill = "none",
    pattern_fill = "none"
  ) +
  labs(
    x = "BMI Category",
    y = "Resting Heart Rate (bpm)"
  ) +
  theme_seasonality()
p_s8_rhr_bmi_box

# LMM for RHR with BMI_CATEGORY * MONTH_LABEL interaction
model_old_rhr_bmi <- lmer(
  RESTING_HEART_RATE ~ BMI_CATEGORY * MONTH_LABEL + PA_TERTILE + GENDER + RHR_person_mean + (1 | USER_ID),
  data = df_old_bmi_aug_dec
)

anova(model_old_rhr_bmi)

emmeans(model_old_rhr_bmi, pairwise ~ BMI_CATEGORY | MONTH_LABEL, adjust = "holm")
emmeans(model_old_rhr_bmi, pairwise ~ MONTH_LABEL | BMI_CATEGORY, adjust = "holm")
```
### HRV Box-And-Whisker and LMM by BMI Category: Old
```{r}
# Reuse df_old_bmi_aug_dec from prior chunk

p_s8_hrv_bmi_box <- ggplot(df_old_bmi_aug_dec, aes(x = BMI_CATEGORY, y = HRV)) +
  geom_boxplot_pattern(
    aes(
      fill = BMI_CATEGORY,
      pattern = MONTH_LABEL,
      pattern_fill = BMI_CATEGORY
    ),
    position = position_dodge(width = 0.75),
    color = "black",
    pattern_density = 0.4,
    pattern_spacing = 0.05,
    alpha = 0.7,
    outlier.shape = NA
  ) +
  stat_summary(
    fun = median,
    geom = "point",
    aes(group = interaction(MONTH_LABEL, BMI_CATEGORY)),
    position = position_dodge(width = 0.75),
    shape = 21,
    size = 3,
    fill = "white",
    color = "black"
  ) +
  scale_fill_manual(values = bmi_colors) +
  scale_pattern_fill_manual(values = bmi_colors) +
  scale_pattern_manual(
    values = c("August" = "none", "December" = "stripe"),
    name = "Month"
  ) +
  guides(
    fill = "none",
    pattern_fill = "none"
  ) +
  labs(
    x = "BMI Category",
    y = "HRV (ms)"
  ) +
  coord_cartesian(ylim = c(0, 150)) +
  theme_seasonality()
p_s8_hrv_bmi_box

model_old_hrv_bmi <- lmer(
  HRV ~ BMI_CATEGORY * MONTH_LABEL + GENDER + PA_TERTILE + HRV_person_mean + (1 | USER_ID),
  data = df_old_bmi_aug_dec
)

anova(model_old_hrv_bmi)
emmeans(model_old_hrv_bmi, pairwise ~ BMI_CATEGORY | MONTH_LABEL, adjust = "holm")
emmeans(model_old_hrv_bmi, pairwise ~ MONTH_LABEL | BMI_CATEGORY, adjust = "holm")

```
### Means+/-SD in BMI categories: Old
```{r}
df_old_bmi %>%
  group_by(BMI_CATEGORY) %>%
  summarise(
    n_users = n_distinct(USER_ID),
    mean_BMI = mean(BMI, na.rm = TRUE),
    sd_BMI = sd(BMI, na.rm = TRUE),
    mean_AGE_2023 = mean(AGE_2023, na.rm = TRUE),
    sd_AGE_2023 = sd(AGE_2023, na.rm = TRUE),
    .groups = "drop"
  )
```

### Middle-Aged Adults: Create Subsets
```{r}
# ---- Create df_mid ONCE (PA analyses) ----
# Inherits whole-sample, user-level PA_TERTILE from df_month (lines 2504-2509)
# so Low/Moderate/High refer to the same absolute activity levels across all
# age strata (df_old, df_mid, df_young).
df_mid <- df_month %>%
  filter(AGE_BUCKET == "Middle")

# ---- Create df_mid_bmi ONCE (BMI analyses) ----
df_mid_bmi <- df_month_bmi %>%
  filter(AGE_BUCKET == "Middle")
```

### RHR Moderation by Activity: Middle
```{r}
# ---- 1. Fit GAMM with cyclic spline by PA_TERTILE ----
formula_rhr_pa_mid <- RHR_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = PA_TERTILE) +
  PA_TERTILE + GENDER + BMI + RHR_person_mean

model_rhr_pa_mid <- bam(
  formula_rhr_pa_mid,
  data = df_mid,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# ---- 2. Marginal trajectory (averages over GENDER; lpmatrix-based SE) ----
v_df_pa <- predict_marginal_trajectory(
  model_rhr_pa_mid,
  data = df_mid,
  by_var = "PA_TERTILE"
)

# ---- 3. Plot (Figure S7: RHR by PA, Middle-Aged) ----
p_s7_rhr_pa <- ggplot(v_df_pa, aes(x = MONTH_NUM, y = visregFit, color = PA_TERTILE, fill = PA_TERTILE)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "Activity Group", values = pa_colors) +
  scale_fill_manual(name = "Activity Group", values = pa_colors) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "RHR Offset (bpm)") +
  theme_seasonality()
p_s7_rhr_pa

# ---- 4. Bootstrap amplitude + pairwise ----
boot_rhr_pa_mid <- bootstrap_amplitude(df_mid, formula_rhr_pa_mid, by = "PA_TERTILE")

bootstrap_summary_pa_mid <- boot_rhr_pa_mid %>%
  group_by(PA_TERTILE) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

boot_pvals_pa_mid <- bootstrap_pairwise(boot_rhr_pa_mid, "PA_TERTILE")
```

### HRV Moderation by Activity: Middle
```{r}
# ---- 1. Fit GAMM with cyclic spline by PA_TERTILE ----
formula_hrv_pa_mid <- HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = PA_TERTILE) +
  PA_TERTILE + GENDER + BMI + HRV_person_mean

model_hrv_pa_mid <- bam(
  formula_hrv_pa_mid,
  data = df_mid,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# ---- 2. Marginal trajectory (averages over GENDER; lpmatrix-based SE) ----
v_df_pa_hrv <- predict_marginal_trajectory(
  model_hrv_pa_mid,
  data = df_mid,
  by_var = "PA_TERTILE"
)

# ---- 3. Plot (Figure S7: HRV by PA, Middle-Aged) ----
p_s7_hrv_pa <- ggplot(v_df_pa_hrv, aes(x = MONTH_NUM, y = visregFit, color = PA_TERTILE, fill = PA_TERTILE)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "Activity Group", values = pa_colors) +
  scale_fill_manual(name = "Activity Group", values = pa_colors) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "HRV Offset (ms)") +
  theme_seasonality()
p_s7_hrv_pa

# ---- 4. Bootstrap amplitude + pairwise ----
boot_hrv_pa_mid <- bootstrap_amplitude(df_mid, formula_hrv_pa_mid, by = "PA_TERTILE")

bootstrap_summary_hrv_pa_mid <- boot_hrv_pa_mid %>%
  group_by(PA_TERTILE) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

boot_pvals_hrv_pa_mid <- bootstrap_pairwise(boot_hrv_pa_mid, "PA_TERTILE")
```

### RHR Moderation by BMI: Middle
```{r}
# ---- 1. Sample sizes by BMI group ----
df_mid_bmi %>%
  group_by(BMI_CATEGORY) %>%
  summarise(n_users = n_distinct(USER_ID), .groups = "drop")

# ---- 2. Fit GAMM with cyclic spline by BMI_CATEGORY ----
formula_rhr_bmi_mid <- RHR_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = BMI_CATEGORY) +
  BMI_CATEGORY + GENDER + SHRZS_b + SHRZS_w + RHR_person_mean

model_rhr_bmi_mid <- bam(
  formula_rhr_bmi_mid,
  data = df_mid_bmi,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# ---- 3. Marginal trajectory (averages over GENDER; lpmatrix-based SE) ----
v_df_bmi_mid <- predict_marginal_trajectory(
  model_rhr_bmi_mid,
  data = df_mid_bmi,
  by_var = "BMI_CATEGORY"
)

# ---- 4. Plot (Figure S7: RHR by BMI, Middle-Aged) ----
p_s7_rhr_bmi <- ggplot(v_df_bmi_mid, aes(x = MONTH_NUM, y = visregFit, color = BMI_CATEGORY, fill = BMI_CATEGORY)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "BMI Group", values = bmi_colors) +
  scale_fill_manual(name = "BMI Group", values = bmi_colors) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "RHR Offset (bpm)") +
  theme_seasonality()
p_s7_rhr_bmi

# ---- 5. Bootstrap amplitude + pairwise (formula matches primary model) ----
boot_rhr_bmi_mid <- bootstrap_amplitude(df_mid_bmi, formula_rhr_bmi_mid, by = "BMI_CATEGORY")

bootstrap_summary_bmi_mid <- boot_rhr_bmi_mid %>%
  group_by(BMI_CATEGORY) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

boot_pvals_bmi_mid <- bootstrap_pairwise(boot_rhr_bmi_mid, "BMI_CATEGORY")
```

### HRV Moderation by BMI: Middle
```{r}
# ---- 1. Fit GAMM with cyclic spline by BMI_CATEGORY ----
formula_hrv_bmi_mid <- HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = BMI_CATEGORY) +
  BMI_CATEGORY + GENDER + SHRZS_b + SHRZS_w + HRV_person_mean

model_hrv_bmi_mid <- bam(
  formula_hrv_bmi_mid,
  data = df_mid_bmi,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# ---- 2. Marginal trajectory (averages over GENDER; lpmatrix-based SE) ----
v_df_bmi_mid <- predict_marginal_trajectory(
  model_hrv_bmi_mid,
  data = df_mid_bmi,
  by_var = "BMI_CATEGORY"
)

# ---- 3. Plot (Figure S7: HRV by BMI, Middle-Aged) ----
p_s7_hrv_bmi <- ggplot(v_df_bmi_mid, aes(x = MONTH_NUM, y = visregFit, color = BMI_CATEGORY, fill = BMI_CATEGORY)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "BMI Group", values = bmi_colors) +
  scale_fill_manual(name = "BMI Group", values = bmi_colors) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "HRV Offset (ms)") +
  theme_seasonality()
p_s7_hrv_bmi

# ---- 4. Bootstrap amplitude + pairwise (formula matches primary model) ----
boot_hrv_bmi_mid <- bootstrap_amplitude(df_mid_bmi, formula_hrv_bmi_mid, by = "BMI_CATEGORY")

bootstrap_summary_hrv_bmi_mid <- boot_hrv_bmi_mid %>%
  group_by(BMI_CATEGORY) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

boot_pvals_hrv_bmi_mid <- bootstrap_pairwise(boot_hrv_bmi_mid, "BMI_CATEGORY")
```


### Young Adults: Create Subsets
```{r}
# ---- Create df_young ONCE (PA analyses) ----
df_young <- df_month %>%
  filter(AGE_BUCKET == "Young")

# ---- Create df_young_bmi ONCE (BMI analyses) ----
df_young_bmi <- df_month_bmi %>%
  filter(AGE_BUCKET == "Young")
```

### RHR Moderation by Activity: Young
```{r}
# ---- 1. Fit GAMM with cyclic spline by PA_TERTILE ----
formula_rhr_pa_young <- RHR_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = PA_TERTILE) +
  PA_TERTILE + GENDER + BMI + RHR_person_mean

model_rhr_pa_y <- bam(
  formula_rhr_pa_young,
  data = df_young,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# ---- 2. Marginal trajectory (averages over GENDER; lpmatrix-based SE) ----
v_df_pa_y <- predict_marginal_trajectory(
  model_rhr_pa_y,
  data = df_young,
  by_var = "PA_TERTILE"
)

# ---- 3. Plot ----
p_s6_rhr_pa <- ggplot(v_df_pa_y, aes(x = MONTH_NUM, y = visregFit, color = PA_TERTILE, fill = PA_TERTILE)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "Activity Group", values = pa_colors) +
  scale_fill_manual(name = "Activity Group", values = pa_colors) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "RHR Offset (bpm)") +
  theme_seasonality()
p_s6_rhr_pa

# ---- 4. Bootstrap amplitude + pairwise ----
boot_rhr_pa_young <- bootstrap_amplitude(df_young, formula_rhr_pa_young, by = "PA_TERTILE")

bootstrap_summary_pa_y <- boot_rhr_pa_young %>%
  group_by(PA_TERTILE) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

boot_pvals_pa_y <- bootstrap_pairwise(boot_rhr_pa_young, "PA_TERTILE")

# ---- 5. Plot Bootstrapped Amplitude by PA_TERTILE ----
bootstrap_summary_pa_y$PA_TERTILE <- factor(bootstrap_summary_pa_y$PA_TERTILE, levels = c("Low", "Moderate", "High"))

p_s6_rhr_pa_amp <- ggplot(bootstrap_summary_pa_y, aes(x = PA_TERTILE, y = mean_amp, fill = PA_TERTILE)) +
  geom_col(width = 0.6, color = "black", alpha = 0.8) +
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high), width = 0.2, linewidth = 1.1) +
  scale_fill_manual(values = pa_colors) +
  labs(
    x = "Physical Activity Tertile (Young Adults)",
    y = "Modeled RHR Amplitude (bpm)"
  ) +
  theme_seasonality(legend = FALSE)
p_s6_rhr_pa_amp
```

### HRV Moderation by Activity: Young
```{r}
# ---- 1. Fit GAMM for HRV with cyclic spline by PA_TERTILE ----
formula_hrv_pa_young <- HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = PA_TERTILE) +
  PA_TERTILE + GENDER + BMI + HRV_person_mean

model_hrv_pa_y <- bam(
  formula_hrv_pa_young,
  data = df_young,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# ---- 2. Marginal trajectory (averages over GENDER; lpmatrix-based SE) ----
v_df_hrv_y <- predict_marginal_trajectory(
  model_hrv_pa_y,
  data = df_young,
  by_var = "PA_TERTILE"
)

# ---- 3. Plot ----
p_s6_hrv_pa <- ggplot(v_df_hrv_y, aes(x = MONTH_NUM, y = visregFit, color = PA_TERTILE, fill = PA_TERTILE)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "Activity Group", values = pa_colors) +
  scale_fill_manual(name = "Activity Group", values = pa_colors) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "HRV Offset (ms)") +
  theme_seasonality()
p_s6_hrv_pa

# ---- 4. Bootstrap amplitude + pairwise ----
boot_hrv_pa_young <- bootstrap_amplitude(df_young, formula_hrv_pa_young, by = "PA_TERTILE")

bootstrap_summary_hrv_y <- boot_hrv_pa_young %>%
  group_by(PA_TERTILE) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

boot_pvals_hrv_y <- bootstrap_pairwise(boot_hrv_pa_young, "PA_TERTILE")
```

### RHR Moderation by BMI: Young
```{r}
# Count participants in each BMI category
df_young_bmi %>%
  group_by(BMI_CATEGORY) %>%
  summarise(n_users = n_distinct(USER_ID), .groups = "drop")

# ---- 1. Fit GAMM with cyclic spline by BMI_CATEGORY ----
formula_rhr_bmi_young <- RHR_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = BMI_CATEGORY) +
  BMI_CATEGORY + GENDER + SHRZS_b + SHRZS_w + RHR_person_mean

model_rhr_bmi_young <- bam(
  formula_rhr_bmi_young,
  data = df_young_bmi,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# ---- 2. Marginal trajectory (averages over GENDER; lpmatrix-based SE) ----
v_df_bmi_young <- predict_marginal_trajectory(
  model_rhr_bmi_young,
  data = df_young_bmi,
  by_var = "BMI_CATEGORY"
)

# ---- 3. Plot ----
p_s6_rhr_bmi <- ggplot(v_df_bmi_young, aes(x = MONTH_NUM, y = visregFit, color = BMI_CATEGORY, fill = BMI_CATEGORY)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "BMI Group", values = bmi_colors) +
  scale_fill_manual(name = "BMI Group", values = bmi_colors) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "RHR Offset (bpm)") +
  theme_seasonality()
p_s6_rhr_bmi

# ---- 4. Bootstrap amplitude + pairwise (formula matches primary model) ----
boot_rhr_bmi_young <- bootstrap_amplitude(df_young_bmi, formula_rhr_bmi_young, by = "BMI_CATEGORY")

bootstrap_summary_bmi_young <- boot_rhr_bmi_young %>%
  group_by(BMI_CATEGORY) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

boot_pvals_bmi_young <- bootstrap_pairwise(boot_rhr_bmi_young, "BMI_CATEGORY")
```

### RHR Moderation by BMI: Young (Plotting Only; No Overweight)
```{r}
df_young_bmi_no <- df_young_bmi %>%
  filter(BMI_CATEGORY %in% c("Normal", "Obese")) %>%
  mutate(BMI_CATEGORY = factor(BMI_CATEGORY, levels = c("Normal", "Obese")))

# Fit GAMM with cyclic spline by BMI_CATEGORY
model_rhr_bmi_young_no <- bam(
  RHR_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = BMI_CATEGORY) +
    BMI_CATEGORY + GENDER + SHRZS_b + SHRZS_w + BMI + RHR_person_mean,
  data = df_young_bmi_no,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# Marginal trajectory (averages over GENDER; lpmatrix-based SE)
v_df_bmi_young_no <- predict_marginal_trajectory(
  model_rhr_bmi_young_no,
  data = df_young_bmi_no,
  by_var = "BMI_CATEGORY"
)

bmi_colors_no <- c("Normal" = "#4daf4a", "Obese" = "#e41a1c")

# Plot
ggplot(v_df_bmi_young_no, aes(x = MONTH_NUM, y = visregFit, color = BMI_CATEGORY, fill = BMI_CATEGORY)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "BMI Group", values = bmi_colors_no) +
  scale_fill_manual(name = "BMI Group", values = bmi_colors_no) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "RHR Offset (bpm)") +
  theme_seasonality()
```

### HRV Moderation by BMI: Young
```{r}
# ---- 1. Fit GAMM with cyclic spline by BMI_CATEGORY ----
formula_hrv_bmi_young <- HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = BMI_CATEGORY) +
  BMI_CATEGORY + GENDER + SHRZS_b + SHRZS_w + HRV_person_mean

model_hrv_bmi_young <- bam(
  formula_hrv_bmi_young,
  data = df_young_bmi,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# ---- 2. Marginal trajectory (averages over GENDER; lpmatrix-based SE) ----
v_df_bmi_hrv_young <- predict_marginal_trajectory(
  model_hrv_bmi_young,
  data = df_young_bmi,
  by_var = "BMI_CATEGORY"
)

# ---- 3. Plot ----
p_s6_hrv_bmi <- ggplot(v_df_bmi_hrv_young, aes(x = MONTH_NUM, y = visregFit, color = BMI_CATEGORY, fill = BMI_CATEGORY)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "BMI Group", values = bmi_colors) +
  scale_fill_manual(name = "BMI Group", values = bmi_colors) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "HRV Offset (ms)") +
  theme_seasonality()
p_s6_hrv_bmi

# ---- 4. Bootstrap amplitude + pairwise (formula matches primary model) ----
boot_hrv_bmi_young <- bootstrap_amplitude(df_young_bmi, formula_hrv_bmi_young, by = "BMI_CATEGORY")

bootstrap_summary_bmi_hrv_young <- boot_hrv_bmi_young %>%
  group_by(BMI_CATEGORY) %>%
  summarise(
    mean_amp = mean(amplitude, na.rm = TRUE),
    ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
    ci_high = quantile(amplitude, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

boot_pvals_bmi_hrv_young <- bootstrap_pairwise(boot_hrv_bmi_young, "BMI_CATEGORY")
```

### HRV Moderation by BMI: Young (Plotting Only; No Overweight)
```{r}
# ---- Fit GAMM with cyclic spline by BMI_CATEGORY (reuses df_young_bmi_no from RHR section) ----
model_hrv_bmi_young_no <- bam(
  HRV_within ~ s(MONTH_NUM, bs = "cc", k = 12, by = BMI_CATEGORY) +
    BMI_CATEGORY + GENDER + BMI + HRV_person_mean,
  data = df_young_bmi_no,
  method = "fREML", knots = MONTH_KNOTS,
  discrete = TRUE
)

# ---- 3. Marginal trajectory (averages over GENDER; lpmatrix-based SE) ----
v_df_hrv_bmi_young_no <- predict_marginal_trajectory(
  model_hrv_bmi_young_no,
  data = df_young_bmi_no,
  by_var = "BMI_CATEGORY"
)

bmi_colors_no <- c("Normal" = "#4daf4a", "Obese" = "#e41a1c")

# ---- 4. Plot ----
ggplot(v_df_hrv_bmi_young_no, aes(x = MONTH_NUM, y = visregFit, color = BMI_CATEGORY, fill = BMI_CATEGORY)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr), alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.8) +
  scale_color_manual(name = "BMI Group", values = bmi_colors_no) +
  scale_fill_manual(name = "BMI Group", values = bmi_colors_no) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  scale_y_continuous(name = "HRV Offset (ms)") +
  theme_seasonality()
```

### RHR Box-And-Whisker and LMM by BMI Category: Young
```{r}
# Subset young adults for Aug/Dec comparison
df_young_bmi_aug_dec <- df_young_bmi %>%
  filter(MONTH_NUM %in% c(8, 12)) %>%
  mutate(
    MONTH_LABEL = factor(ifelse(MONTH_NUM == 8, "August", "December"), levels = c("August", "December")),
    pattern = ifelse(MONTH_LABEL == "December", "stripe", "none")
  )

# Plot
p_s6_rhr_box <- ggplot(df_young_bmi_aug_dec, aes(x = BMI_CATEGORY, y = RESTING_HEART_RATE)) +
  geom_boxplot_pattern(
    aes(
      fill = BMI_CATEGORY,
      pattern = MONTH_LABEL,
      pattern_fill = BMI_CATEGORY
    ),
    position = position_dodge(width = 0.75),
    color = "black",
    pattern_density = 0.4,
    pattern_spacing = 0.05,
    alpha = 0.7,
    outlier.shape = NA
  ) +
  stat_summary(
    fun = median,
    geom = "point",
    aes(group = interaction(MONTH_LABEL, BMI_CATEGORY)),
    position = position_dodge(width = 0.75),
    shape = 21,
    size = 3,
    fill = "white",
    color = "black"
  ) +
  scale_fill_manual(values = bmi_colors) +
  scale_pattern_fill_manual(values = bmi_colors) +
  scale_pattern_manual(
    values = c("August" = "none", "December" = "stripe"),
    name = "Month"
  ) +
  guides(
    fill = "none",
    pattern_fill = "none"
  ) +
  labs(
    x = "BMI Category",
    y = "Resting Heart Rate (bpm)"
  ) +
  theme_seasonality()
p_s6_rhr_box

# Linear Mixed Model
model_young_rhr_bmi <- lmer(
  RESTING_HEART_RATE ~ BMI_CATEGORY * MONTH_LABEL + GENDER + PA_TERTILE + RHR_person_mean + (1 | USER_ID),
  data = df_young_bmi_aug_dec
)

# Model ANOVA
anova(model_young_rhr_bmi)

# BMI category differences within each month
emmeans(model_young_rhr_bmi, pairwise ~ BMI_CATEGORY | MONTH_LABEL, adjust = "holm")

# Month differences within each BMI category
emmeans(model_young_rhr_bmi, pairwise ~ MONTH_LABEL | BMI_CATEGORY, adjust = "holm")
```

### HRV Box-And-Whisker and LMM by BMI Category: Young
```{r}
# Plot
p_s6_hrv_box <- ggplot(df_young_bmi_aug_dec, aes(x = BMI_CATEGORY, y = HRV)) +
  geom_boxplot_pattern(
    aes(
      fill = BMI_CATEGORY,
      pattern = MONTH_LABEL,
      pattern_fill = BMI_CATEGORY
    ),
    position = position_dodge(width = 0.75),
    color = "black",
    pattern_density = 0.4,
    pattern_spacing = 0.05,
    alpha = 0.7,
    outlier.shape = NA
  ) +
  stat_summary(
    fun = median,
    geom = "point",
    aes(group = interaction(MONTH_LABEL, BMI_CATEGORY)),
    position = position_dodge(width = 0.75),
    shape = 21,
    size = 3,
    fill = "white",
    color = "black"
  ) +
  scale_fill_manual(values = bmi_colors) +
  scale_pattern_fill_manual(values = bmi_colors) +
  scale_pattern_manual(
    values = c("August" = "none", "December" = "stripe"),
    name = "Month"
  ) +
  guides(
    fill = "none",
    pattern_fill = "none"
  ) +
  labs(
    x = "BMI Category",
    y = "HRV (ms)"
  ) +
  coord_cartesian(ylim = c(0, 150)) +
  theme_seasonality()
p_s6_hrv_box

# Linear Mixed Model for HRV (same structure as RHR model)
model_young_hrv_bmi <- lmer(
  HRV ~ BMI_CATEGORY * MONTH_LABEL + GENDER + PA_TERTILE + HRV_person_mean + (1 | USER_ID),
  data = df_young_bmi_aug_dec
)

# Type III ANOVA for fixed effects
anova(model_young_hrv_bmi)

emmeans(model_young_hrv_bmi, pairwise ~ BMI_CATEGORY | MONTH_LABEL, adjust = "holm")
emmeans(model_young_hrv_bmi, pairwise ~ MONTH_LABEL | BMI_CATEGORY, adjust = "holm")
```

### Means+/-SD in BMI categories: Young
```{r}
df_young_bmi %>%
  group_by(BMI_CATEGORY) %>%
  summarise(
    n_users = n_distinct(USER_ID),
    mean_BMI = mean(BMI, na.rm = TRUE),
    sd_BMI = sd(BMI, na.rm = TRUE),
    mean_AGE_2023 = mean(AGE_2023, na.rm = TRUE),
    sd_AGE_2023 = sd(AGE_2023, na.rm = TRUE),
    .groups = "drop"
  )
```


### Monthly summary of physical activity and sleep
```{r}
df_behavior_monthly <- df_month %>%
  group_by(MONTH_NUM) %>%
  summarise(
    Sleep = mean(MINUTES_OF_SLEEP, na.rm = TRUE),
    SHRZS = mean(SHRZS, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(-MONTH_NUM, names_to = "Behavior", values_to = "Value")

ggplot(df_behavior_monthly, aes(x = MONTH_NUM, y = Value)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2) +
  facet_wrap(~ Behavior, scales = "free_y", ncol = 1,
             labeller = as_labeller(c(
               Sleep = "Sleep Duration (min/night)",
               SHRZS = "Physical Activity (SHRZS)"
             ))) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  theme_seasonality(legend = FALSE) +
  theme(strip.text = element_text(face = "bold"))
```

### Behavioral Seasonal Trajectories (Figures 3a–c)
```{r}
# 1. Fit cyclic GAMs
model_sleep <- gam(MINUTES_OF_SLEEP ~ s(MONTH_NUM, bs = "cc", k = 12),
                   data = df_month, method = "REML", knots = MONTH_KNOTS)

model_activity <- gam(SHRZS ~ s(MONTH_NUM, bs = "cc", k = 12),
                      data = df_month, method = "REML", knots = MONTH_KNOTS)

# 2. Extract smooth fits with visreg
vis_sleep <- visreg(model_sleep, xvar = "MONTH_NUM", plot = FALSE)
vis_activity <- visreg(model_activity, xvar = "MONTH_NUM", plot = FALSE)

# 3. Build tidy data frames for plotting
df_sleep <- vis_sleep$fit %>%
  mutate(Behavior = "Sleep Duration (min/night)")

df_activity <- vis_activity$fit %>%
  mutate(Behavior = "Physical Activity (SHRZS)")

# 4. Define custom colors
behavior_colors <- c(
  "Sleep Duration (min/night)" = "#fc8d62",     # Orange
  "Physical Activity (SHRZS)" = "#1f78b4"       # Blue
)

# 5. Plot each behavior separately (Figure 3a-b)

# --- Sleep (Figure 3b) ---
p_fig3b <- ggplot(df_sleep, aes(x = MONTH_NUM, y = visregFit)) +
  geom_line(color = behavior_colors["Sleep Duration (min/night)"], linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr),
              fill = behavior_colors["Sleep Duration (min/night)"], alpha = 0.3) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  ylab("Sleep Duration (min/night)") +
  theme_seasonality(legend = FALSE)

# --- Activity (Figure 3a) ---
p_fig3a <- ggplot(df_activity, aes(x = MONTH_NUM, y = visregFit)) +
  geom_line(color = behavior_colors["Physical Activity (SHRZS)"], linewidth = 1.2) +
  geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr),
              fill = behavior_colors["Physical Activity (SHRZS)"], alpha = 0.3) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
  ylab("Physical Activity (SHRZS)") +
  theme_seasonality(legend = FALSE)

p_fig3a
p_fig3b

get_peak_nadir <- function(df, behavior_label) {
  df <- df %>% mutate(Behavior = behavior_label)

  peak <- df[which.max(df$visregFit), ]
  nadir <- df[which.min(df$visregFit), ]

  bind_rows(
    peak = peak,
    nadir = nadir,
    .id = "Extreme"
  )
}

extremes_sleep <- get_peak_nadir(df_sleep, "Sleep Duration (min/night)")
extremes_activity <- get_peak_nadir(df_activity, "Physical Activity (SHRZS)")

all_extremes <- bind_rows(
  extremes_sleep,
  extremes_activity
)

# Add readable month name
all_extremes <- all_extremes %>%
  mutate(Month = month.abb[MONTH_NUM]) %>%
  select(Behavior, Extreme, MONTH_NUM, Month, visregFit)
```

```{r}
# Add AGE_BUCKET as factor to ensure consistent ordering
df_month <- df_month %>%
  mutate(AGE_BUCKET = factor(AGE_BUCKET, levels = c("Young", "Middle", "Old")))

# 1. Fit cyclic models by AGE_BUCKET
model_sleep <- gam(
  MINUTES_OF_SLEEP ~ s(MONTH_NUM, bs = "cc", k = 12, by = AGE_BUCKET) + AGE_BUCKET,
  data = df_month, method = "REML", knots = MONTH_KNOTS
)

model_activity <- gam(
  SHRZS ~ s(MONTH_NUM, bs = "cc", k = 12, by = AGE_BUCKET) + AGE_BUCKET,
  data = df_month, method = "REML", knots = MONTH_KNOTS
)

vis_sleep <- visreg(model_sleep, xvar = "MONTH_NUM", by = "AGE_BUCKET",
                    overlay = FALSE, plot = FALSE)

vis_activity <- visreg(model_activity, xvar = "MONTH_NUM", by = "AGE_BUCKET",
                       overlay = FALSE, plot = FALSE)

df_sleep_age <- vis_sleep$fit %>%
  mutate(Behavior = "Sleep Duration (min/night)")

df_activity_age <- vis_activity$fit %>%
  mutate(Behavior = "Physical Activity (SHRZS)")

age_linetypes <- c(
  "Young" = "solid",
  "Middle" = "dashed",
  "Old" = "dotted"
)

plot_behavior <- function(df_behavior, color) {
  ggplot(df_behavior, aes(x = MONTH_NUM, y = visregFit,
                          linetype = AGE_BUCKET)) +
    geom_line(color = color, linewidth = 1.2) +
    geom_ribbon(aes(ymin = visregLwr, ymax = visregUpr),
                fill = color, alpha = 0.15, color = NA) +
    scale_linetype_manual(values = age_linetypes) +
    scale_x_continuous(breaks = 1:12, labels = month.abb, name = NULL) +
    labs(x = "Month",
         y = unique(df_behavior$Behavior)) +
    theme_seasonality(legend = FALSE)
}

# Figure S4: age-stratified behavioral seasonal curves
p_s4b <- plot_behavior(df_sleep_age,
                       behavior_colors["Sleep Duration (min/night)"])

p_s4a <- plot_behavior(df_activity_age,
                       behavior_colors["Physical Activity (SHRZS)"])

p_s4a
p_s4b

legend_df <- data.frame(
  AGE_BUCKET = c("Young", "Middle", "Old"),
  x = c(1, 4, 7),
  xend = c(2.5, 5.5, 8.5),
  y = 1,
  Label = c("Young", "Middle", "Old")
)

age_bucket_legend_horizontal <- ggplot(legend_df) +
  geom_segment(
    aes(x = x, xend = xend, y = y, yend = y, linetype = AGE_BUCKET),
    color = "black", linewidth = 1.2
  ) +
  geom_text(
    aes(x = xend + 0.3, y = y, label = Label),
    hjust = 0, vjust = 0.5, fontface = "bold", size = 5
  ) +
  scale_linetype_manual(values = age_linetypes) +
  coord_cartesian(xlim = c(0.5, 10), ylim = c(0.5, 1.5), clip = "off") +
  theme_void() +
  theme(
    legend.position = "none",
    plot.margin = margin(0, 0, 0, 0)
  )

age_bucket_legend_horizontal
```

### Assemble Multi-Panel Figures
```{r, fig.width=14, fig.height=12}
# Assembles Figures 1-4 from component plots created above and writes PDF+PNG to the Main Figures directory.
fig_dir <- "/Users/gregory.grosicki/Desktop/Seasonality/Main Figures"
sup_dir <- "/Users/gregory.grosicki/Desktop/Seasonality/Supplementary Figures"

# ---- Figure 1: RHR and HRV by Age Group (2x2) ----
fig1 <- ((p_fig1a + p_fig1b) / (p_fig1c + p_fig1d)) +
  plot_annotation(tag_levels = "A") +
  theme(plot.tag = element_text(size = 16, face = "bold"))
fig1
ggsave(file.path(fig_dir, "Figure_1.pdf"), fig1, width = 14, height = 10, dpi = 300)
ggsave(file.path(fig_dir, "Figure_1.png"), fig1, width = 14, height = 10, dpi = 300)

# ---- Figure 2: RHR and HRV by Sex (2x2) ----
fig2 <- ((p_fig2a + p_fig2b) / (p_fig2c + p_fig2d)) +
  plot_annotation(tag_levels = "A") +
  theme(plot.tag = element_text(size = 16, face = "bold"))
fig2
ggsave(file.path(fig_dir, "Figure_2.pdf"), fig2, width = 14, height = 10, dpi = 300)
ggsave(file.path(fig_dir, "Figure_2.png"), fig2, width = 14, height = 10, dpi = 300)

# ---- Figure 3: Behavioral Attenuation (3x3) ----
fig3_theme <- theme(axis.text = element_text(size = 12, face = "bold"),
                    axis.title = element_text(size = 14, face = "bold"))

fig3 <- ((p_fig3a + fig3_theme) + (p_fig3b + fig3_theme)) /
        ((p_fig3d + fig3_theme) + (p_fig3e + fig3_theme)) /
        ((p_fig3g + fig3_theme) + (p_fig3h + fig3_theme)) +
  plot_annotation(tag_levels = "A") +
  theme(plot.tag = element_text(size = 18, face = "bold"))
fig3
ggsave(file.path(fig_dir, "Figure_3.pdf"), fig3, width = 16, height = 14, dpi = 300)
ggsave(file.path(fig_dir, "Figure_3.png"), fig3, width = 16, height = 14, dpi = 300)

# ---- Figure 4: PA Level and BMI Moderation (2x2) ----
fig4 <- ((p_fig4a + p_fig4b) / (p_fig4c + p_fig4d)) +
  plot_annotation(tag_levels = "A") +
  theme(plot.tag = element_text(size = 16, face = "bold"))
fig4
ggsave(file.path(fig_dir, "Figure_4.pdf"), fig4, width = 14, height = 10, dpi = 300)
ggsave(file.path(fig_dir, "Figure_4.png"), fig4, width = 14, height = 10, dpi = 300)
```

### Supplementary Figure Assembly (Figures S1–S9)
```{r, fig.width=14, fig.height=12}
# ---- Figure S1: Overall RHR and HRV (1x2) ----
fig_s1 <- (p_s1a + p_s1b) +
  plot_annotation(tag_levels = "A") +
  theme(plot.tag = element_text(size = 16, face = "bold"))
fig_s1
ggsave(file.path(sup_dir, "Figure_S1.pdf"), fig_s1, width = 14, height = 5, dpi = 300)
ggsave(file.path(sup_dir, "Figure_S1.png"), fig_s1, width = 14, height = 5, dpi = 300)

# ---- Figure S2: RHR and HRV by Age – GAMM + Box-Whisker (2x2) ----
fig_s2 <- ((p_s2a + p_s2b) / (p_s2c + p_s2d)) +
  plot_annotation(tag_levels = "A") +
  theme(plot.tag = element_text(size = 16, face = "bold"))
fig_s2
ggsave(file.path(sup_dir, "Figure_S2.pdf"), fig_s2, width = 14, height = 10, dpi = 300)
ggsave(file.path(sup_dir, "Figure_S2.png"), fig_s2, width = 14, height = 10, dpi = 300)

# ---- Figure S3: RHR and HRV by Sex – GAMM + Box-Whisker (2x2) ----
fig_s3 <- ((p_s3a + p_s3b) / (p_s3c + p_s3d)) +
  plot_annotation(tag_levels = "A") +
  theme(plot.tag = element_text(size = 16, face = "bold"))
fig_s3
ggsave(file.path(sup_dir, "Figure_S3.pdf"), fig_s3, width = 14, height = 10, dpi = 300)
ggsave(file.path(sup_dir, "Figure_S3.png"), fig_s3, width = 14, height = 10, dpi = 300)

# ---- Figure S4: Behavioral Patterns by Age Group (1x2) ----
fig_s4 <- (p_s4a + p_s4b + plot_layout(ncol = 2)) +
  plot_annotation(tag_levels = "A") +
  theme(plot.tag = element_text(size = 16, face = "bold"))
fig_s4
ggsave(file.path(sup_dir, "Figure_S4.pdf"), fig_s4, width = 12, height = 5, dpi = 300)
ggsave(file.path(sup_dir, "Figure_S4.png"), fig_s4, width = 12, height = 5, dpi = 300)

# NOTE: After integrating photoperiod into the behavioral-mediation narrative,
# photoperiod (previously S9) becomes the new Figure S5 (appears first in text),
# age-stratified all-adjusted becomes the new Figure S6, and the Aim-3 age-stratified
# moderation figures are S7/S8/S9. Internal R object names (fig_s5, fig_s6, fig_s7,
# fig_s8, fig_s9) reflect original creation order; saved filenames reflect final
# numbering.

# ---- Figure S5 (new): RHR and HRV amplitude vs annual daylight amplitude (1x2) ----
fig_s9 <- (p_s9a + p_s9b) +
  plot_annotation(tag_levels = "A") +
  theme(plot.tag = element_text(size = 16, face = "bold"))
fig_s9
ggsave(file.path(sup_dir, "Figure_S5.pdf"), fig_s9, width = 14, height = 5, dpi = 300)
ggsave(file.path(sup_dir, "Figure_S5.png"), fig_s9, width = 14, height = 5, dpi = 300)

# ---- Figure S6 (new): Adjusted RHR and HRV by Age (1x2) ----
fig_s5 <- (p_s5a + p_s5b) +
  plot_annotation(tag_levels = "A") +
  theme(plot.tag = element_text(size = 16, face = "bold"))
fig_s5
ggsave(file.path(sup_dir, "Figure_S6.pdf"), fig_s5, width = 14, height = 5, dpi = 300)
ggsave(file.path(sup_dir, "Figure_S6.png"), fig_s5, width = 14, height = 5, dpi = 300)

# ---- Figure S7 (new): Young Adults – PA and BMI (2x2) ----
fig_s6 <- ((p_s6_rhr_pa + p_s6_hrv_pa) / (p_s6_rhr_bmi + p_s6_hrv_bmi)) +
  plot_annotation(tag_levels = "A") +
  theme(plot.tag = element_text(size = 16, face = "bold"))
fig_s6
ggsave(file.path(sup_dir, "Figure_S7.pdf"), fig_s6, width = 14, height = 10, dpi = 300)
ggsave(file.path(sup_dir, "Figure_S7.png"), fig_s6, width = 14, height = 10, dpi = 300)

# ---- Figure S8 (new): Middle-Aged Adults – PA and BMI (2x2) ----
fig_s7 <- ((p_s7_rhr_pa + p_s7_hrv_pa) / (p_s7_rhr_bmi + p_s7_hrv_bmi)) +
  plot_annotation(tag_levels = "A") +
  theme(plot.tag = element_text(size = 16, face = "bold"))
fig_s7
ggsave(file.path(sup_dir, "Figure_S8.pdf"), fig_s7, width = 14, height = 10, dpi = 300)
ggsave(file.path(sup_dir, "Figure_S8.png"), fig_s7, width = 14, height = 10, dpi = 300)

# ---- Figure S9 (new): Older Adults – PA and BMI (2x2) ----
fig_s8 <- ((p_s8_rhr_pa + p_s8_hrv_pa) / (p_s8_rhr_bmi + p_s8_hrv_bmi)) +
  plot_annotation(tag_levels = "A") +
  theme(plot.tag = element_text(size = 16, face = "bold"))
fig_s8
ggsave(file.path(sup_dir, "Figure_S9.pdf"), fig_s8, width = 14, height = 10, dpi = 300)
ggsave(file.path(sup_dir, "Figure_S9.png"), fig_s8, width = 14, height = 10, dpi = 300)
```

### Export Results as CSVs
```{r}
# Exports bootstrap amplitude summaries, pairwise p-values, attenuation estimates, and Aug/Dec contrasts for every aim to the Results CSVs directory. Each numbered block below corresponds to a figure in the paper.
csv_dir <- "/Users/gregory.grosicki/Desktop/Seasonality/Results CSVs"
dir.create(csv_dir, showWarnings = FALSE, recursive = TRUE)

# ---- 1. Overall amplitude (Figure S1) ----
overall_amp <- tibble(
  Outcome = c("RHR", "HRV"),
  mean_amp = c(
    mean(boot_rhr_overall$amplitude, na.rm = TRUE),
    mean(boot_hrv_overall$amplitude, na.rm = TRUE)
  ),
  ci_low = c(
    quantile(boot_rhr_overall$amplitude, 0.025, na.rm = TRUE),
    quantile(boot_hrv_overall$amplitude, 0.025, na.rm = TRUE)
  ),
  ci_high = c(
    quantile(boot_rhr_overall$amplitude, 0.975, na.rm = TRUE),
    quantile(boot_hrv_overall$amplitude, 0.975, na.rm = TRUE)
  )
)
write.csv(overall_amp, file.path(csv_dir, "overall_amplitude.csv"), row.names = FALSE)

# ---- 2. Amplitude by Age – GAM bootstrap (Figure 1A/C) ----
boot_age_rhr <- boot_rhr_age %>%
  group_by(AGE_BUCKET) %>%
  summarise(mean_amp = mean(amplitude, na.rm = TRUE),
            ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
            ci_high = quantile(amplitude, 0.975, na.rm = TRUE), .groups = "drop") %>%
  mutate(Outcome = "RHR")
boot_age_hrv <- boot_hrv_age %>%
  group_by(AGE_BUCKET) %>%
  summarise(mean_amp = mean(amplitude, na.rm = TRUE),
            ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
            ci_high = quantile(amplitude, 0.975, na.rm = TRUE), .groups = "drop") %>%
  mutate(Outcome = "HRV")
write.csv(bind_rows(boot_age_rhr, boot_age_hrv), file.path(csv_dir, "amplitude_by_age_GAM.csv"), row.names = FALSE)

# Pairwise p-values
write.csv(
  bind_rows(boot_pvals %>% mutate(Outcome = "RHR"), boot_pvals_hrv_age %>% mutate(Outcome = "HRV")),
  file.path(csv_dir, "amplitude_by_age_GAM_pvals.csv"), row.names = FALSE
)

# ---- 3. Amplitude by Age – GAMM point estimates (Figure S2A/C) ----
write.csv(
  bind_rows(amplitude_gamm %>% mutate(Outcome = "RHR"), amplitude_gamm_hrv %>% mutate(Outcome = "HRV")),
  file.path(csv_dir, "amplitude_by_age_GAMM.csv"), row.names = FALSE
)

# ---- 4. GAMM pairwise contrasts by Age at Aug/Dec (Figure S2B/D) ----
write.csv(contrasts_all %>% mutate(Outcome = "RHR"), file.path(csv_dir, "GAMM_age_contrasts_RHR.csv"), row.names = FALSE)
# HRV contrasts (same structure computed in HRV GAMM section)
if (exists("contrasts_all_hrv")) {
  write.csv(contrasts_all_hrv %>% mutate(Outcome = "HRV"), file.path(csv_dir, "GAMM_age_contrasts_HRV.csv"), row.names = FALSE)
}

# ---- 5. Individual-level amplitude by Age (Figure 1B/D) ----
write.csv(amplitude_summary %>% mutate(Outcome = "RHR"), file.path(csv_dir, "individual_amplitude_by_age_RHR.csv"), row.names = FALSE)
if (exists("amplitude_summary_hrv")) {
  write.csv(amplitude_summary_hrv %>% mutate(Outcome = "HRV"), file.path(csv_dir, "individual_amplitude_by_age_HRV.csv"), row.names = FALSE)
}

# ---- 6. Amplitude by Sex – GAM bootstrap (Figure 2A/C) ----
boot_sex_rhr <- boot_rhr_sex %>%
  group_by(GENDER) %>%
  summarise(mean_amp = mean(amplitude, na.rm = TRUE),
            ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
            ci_high = quantile(amplitude, 0.975, na.rm = TRUE), .groups = "drop") %>%
  mutate(Outcome = "RHR")
boot_sex_hrv <- boot_hrv_sex %>%
  group_by(GENDER) %>%
  summarise(mean_amp = mean(amplitude, na.rm = TRUE),
            ci_low = quantile(amplitude, 0.025, na.rm = TRUE),
            ci_high = quantile(amplitude, 0.975, na.rm = TRUE), .groups = "drop") %>%
  mutate(Outcome = "HRV")
write.csv(bind_rows(boot_sex_rhr, boot_sex_hrv), file.path(csv_dir, "amplitude_by_sex_GAM.csv"), row.names = FALSE)

# Pairwise p-values
write.csv(
  bind_rows(boot_pvals_sex_rhr %>% mutate(Outcome = "RHR"), boot_pvals_sex_hrv %>% mutate(Outcome = "HRV")),
  file.path(csv_dir, "amplitude_by_sex_GAM_pvals.csv"), row.names = FALSE
)

# ---- 7. Amplitude by Sex – GAMM point estimates (Figure S3A/C) ----
write.csv(
  bind_rows(amplitude_gamm_gender %>% mutate(Outcome = "RHR"), amplitude_gamm_gender_hrv %>% mutate(Outcome = "HRV")),
  file.path(csv_dir, "amplitude_by_sex_GAMM.csv"), row.names = FALSE
)

# ---- 8. GAMM pairwise contrasts by Sex at Aug/Dec (Figure S3B/D) ----
write.csv(gender_contrasts %>% mutate(Outcome = "RHR"), file.path(csv_dir, "GAMM_sex_contrasts_RHR.csv"), row.names = FALSE)
write.csv(gender_contrasts_hrv %>% mutate(Outcome = "HRV"), file.path(csv_dir, "GAMM_sex_contrasts_HRV.csv"), row.names = FALSE)

# ---- 9. Attenuation results (Figure 3D-I) ----
extract_attenuation <- function(att, label) {
  amp_diff <- att$amp_base - att$amp_adj
  tibble(
    Covariate = label,
    mean_base = mean(att$amp_base),
    ci_base_low = quantile(att$amp_base, 0.025),
    ci_base_high = quantile(att$amp_base, 0.975),
    mean_adj = mean(att$amp_adj),
    ci_adj_low = quantile(att$amp_adj, 0.025),
    ci_adj_high = quantile(att$amp_adj, 0.975),
    mean_diff = mean(att$amp_base) - mean(att$amp_adj),
    ci_diff_low = quantile(amp_diff, 0.025),
    ci_diff_high = quantile(amp_diff, 0.975),
    pct_attenuation = (mean(att$amp_base) - mean(att$amp_adj)) / mean(att$amp_base) * 100,
    n_valid = att$n_valid
  )
}

attenuation_results <- bind_rows(
  extract_attenuation(att_rhr_pa, "PA") %>% mutate(Outcome = "RHR"),
  extract_attenuation(att_hrv_pa, "PA") %>% mutate(Outcome = "HRV"),
  extract_attenuation(att_rhr_sleep, "Sleep") %>% mutate(Outcome = "RHR"),
  extract_attenuation(att_hrv_sleep, "Sleep") %>% mutate(Outcome = "HRV")
)
write.csv(attenuation_results, file.path(csv_dir, "attenuation_results.csv"), row.names = FALSE)

# ---- 10. Adjusted amplitude by Age (Figure S5) ----
write.csv(
  bind_rows(bootstrap_summary_rhr %>% mutate(Outcome = "RHR"), bootstrap_summary_hrv_adj %>% mutate(Outcome = "HRV")),
  file.path(csv_dir, "adjusted_amplitude_by_age.csv"), row.names = FALSE
)
write.csv(
  bind_rows(boot_pvals_rhr %>% mutate(Outcome = "RHR"), boot_pvals_hrv_adj %>% mutate(Outcome = "HRV")),
  file.path(csv_dir, "adjusted_amplitude_by_age_pvals.csv"), row.names = FALSE
)

# ---- 11. PA moderation – All ages (Figure 4A/B) ----
write.csv(
  bind_rows(bootstrap_summary_pa_all %>% mutate(Outcome = "RHR"), bootstrap_summary_hrv_pa_all %>% mutate(Outcome = "HRV")),
  file.path(csv_dir, "amplitude_by_PA_all.csv"), row.names = FALSE
)
write.csv(
  bind_rows(boot_pvals_pa_all %>% mutate(Outcome = "RHR"), boot_pvals_hrv_pa_all %>% mutate(Outcome = "HRV")),
  file.path(csv_dir, "amplitude_by_PA_all_pvals.csv"), row.names = FALSE
)

# ---- 12. BMI moderation – All ages (Figure 4C/D) ----
write.csv(
  bind_rows(bootstrap_summary_bmi_all %>% mutate(Outcome = "RHR"), bootstrap_summary_hrv_bmi_all %>% mutate(Outcome = "HRV")),
  file.path(csv_dir, "amplitude_by_BMI_all.csv"), row.names = FALSE
)
write.csv(
  bind_rows(boot_pvals_bmi_all %>% mutate(Outcome = "RHR"), boot_pvals_hrv_bmi_all %>% mutate(Outcome = "HRV")),
  file.path(csv_dir, "amplitude_by_BMI_all_pvals.csv"), row.names = FALSE
)

# ---- 13. Age-stratified PA moderation (Figures S6-S8) ----
# Young
write.csv(
  bind_rows(bootstrap_summary_pa_y %>% mutate(Outcome = "RHR"), bootstrap_summary_hrv_y %>% mutate(Outcome = "HRV")),
  file.path(csv_dir, "amplitude_by_PA_young.csv"), row.names = FALSE
)
write.csv(
  bind_rows(boot_pvals_pa_y %>% mutate(Outcome = "RHR"), boot_pvals_hrv_y %>% mutate(Outcome = "HRV")),
  file.path(csv_dir, "amplitude_by_PA_young_pvals.csv"), row.names = FALSE
)

# Middle
write.csv(
  bind_rows(bootstrap_summary_pa_mid %>% mutate(Outcome = "RHR"),
            bootstrap_summary_hrv_pa_mid %>% mutate(Outcome = "HRV")),
  file.path(csv_dir, "amplitude_by_PA_middle.csv"), row.names = FALSE
)
write.csv(
  bind_rows(boot_pvals_pa_mid %>% mutate(Outcome = "RHR"),
            boot_pvals_hrv_pa_mid %>% mutate(Outcome = "HRV")),
  file.path(csv_dir, "amplitude_by_PA_middle_pvals.csv"), row.names = FALSE
)

# Old
write.csv(
  bind_rows(bootstrap_summary_pa_old %>% mutate(Outcome = "RHR"),
            bootstrap_summary_hrv_pa_old %>% mutate(Outcome = "HRV")),
  file.path(csv_dir, "amplitude_by_PA_old.csv"), row.names = FALSE
)
write.csv(
  bind_rows(boot_pvals_pa_old %>% mutate(Outcome = "RHR"),
            boot_pvals_hrv_pa_old %>% mutate(Outcome = "HRV")),
  file.path(csv_dir, "amplitude_by_PA_old_pvals.csv"), row.names = FALSE
)

# ---- 14. Age-stratified BMI moderation (Figures S6-S8) ----
# Young
write.csv(
  bind_rows(bootstrap_summary_bmi_young %>% mutate(Outcome = "RHR"),
            bootstrap_summary_bmi_hrv_young %>% mutate(Outcome = "HRV")),
  file.path(csv_dir, "amplitude_by_BMI_young.csv"), row.names = FALSE
)
write.csv(
  bind_rows(boot_pvals_bmi_young %>% mutate(Outcome = "RHR"),
            boot_pvals_bmi_hrv_young %>% mutate(Outcome = "HRV")),
  file.path(csv_dir, "amplitude_by_BMI_young_pvals.csv"), row.names = FALSE
)

# Middle
write.csv(
  bind_rows(bootstrap_summary_bmi_mid %>% mutate(Outcome = "RHR"),
            bootstrap_summary_hrv_bmi_mid %>% mutate(Outcome = "HRV")),
  file.path(csv_dir, "amplitude_by_BMI_middle.csv"), row.names = FALSE
)
write.csv(
  bind_rows(boot_pvals_bmi_mid %>% mutate(Outcome = "RHR"),
            boot_pvals_hrv_bmi_mid %>% mutate(Outcome = "HRV")),
  file.path(csv_dir, "amplitude_by_BMI_middle_pvals.csv"), row.names = FALSE
)

# Old
write.csv(
  bind_rows(bootstrap_summary_bmi_old %>% mutate(Outcome = "RHR"),
            bootstrap_summary_hrv_bmi_old %>% mutate(Outcome = "HRV")),
  file.path(csv_dir, "amplitude_by_BMI_old.csv"), row.names = FALSE
)
write.csv(
  bind_rows(boot_pvals_bmi_old %>% mutate(Outcome = "RHR"),
            boot_pvals_hrv_bmi_old %>% mutate(Outcome = "HRV")),
  file.path(csv_dir, "amplitude_by_BMI_old_pvals.csv"), row.names = FALSE
)

# ---- 15. Mixed model ANOVA and emmeans for Aug/Dec comparisons ----
# Age × Month (RHR)
emm_age_month_rhr <- as.data.frame(emmeans(model_aug_dec_rhr_age, pairwise ~ AGE_BUCKET | MONTH_LABEL, adjust = "holm")$contrasts)
emm_month_age_rhr <- as.data.frame(emmeans(model_aug_dec_rhr_age, pairwise ~ MONTH_LABEL | AGE_BUCKET, adjust = "holm")$contrasts)
write.csv(emm_age_month_rhr, file.path(csv_dir, "emmeans_age_within_month_RHR.csv"), row.names = FALSE)
write.csv(emm_month_age_rhr, file.path(csv_dir, "emmeans_month_within_age_RHR.csv"), row.names = FALSE)

# Age × Month (HRV)
emm_age_month_hrv <- as.data.frame(emmeans(model_aug_dec_hrv_age, pairwise ~ AGE_BUCKET | MONTH_LABEL, adjust = "holm")$contrasts)
emm_month_age_hrv <- as.data.frame(emmeans(model_aug_dec_hrv_age, pairwise ~ MONTH_LABEL | AGE_BUCKET, adjust = "holm")$contrasts)
write.csv(emm_age_month_hrv, file.path(csv_dir, "emmeans_age_within_month_HRV.csv"), row.names = FALSE)
write.csv(emm_month_age_hrv, file.path(csv_dir, "emmeans_month_within_age_HRV.csv"), row.names = FALSE)

# Sex × Month (RHR)
anova_sex_rhr <- as.data.frame(anova(model_aug_dec_rhr_gender))
write.csv(anova_sex_rhr, file.path(csv_dir, "anova_sex_month_RHR.csv"), row.names = TRUE)

# Sex × Month (HRV)
anova_sex_hrv <- as.data.frame(anova(model_hrv_gender_interact))
write.csv(anova_sex_hrv, file.path(csv_dir, "anova_sex_month_HRV.csv"), row.names = TRUE)

# Individual-level amplitude Tukey (HRV)
emm_hrv_age <- emmeans(model_amplitude_hrv, ~ AGE_BUCKET)
tukey_hrv_age <- as.data.frame(pairs(emm_hrv_age, adjust = "tukey"))
write.csv(tukey_hrv_age, file.path(csv_dir, "individual_amplitude_tukey_HRV.csv"), row.names = FALSE)

# Individual-level amplitude by Sex (RHR)
  tukey_sex_rhr <- as.data.frame(pairs(emmeans(model_amplitude_gender, ~ GENDER), adjust = "tukey"))                                                                               
  write.csv(tukey_sex_rhr, file.path(csv_dir, "individual_amplitude_tukey_sex_RHR.csv"), row.names = FALSE)   
  
  tukey_sex_hrv <- as.data.frame(pairs(emmeans(model_amplitude_gender_hrv, ~ GENDER), adjust = "tukey"))
  write.csv(tukey_sex_hrv, file.path(csv_dir, "individual_amplitude_tukey_sex_HRV.csv"), row.names = FALSE)

# ---- 16. Geographic daylight-amplitude regression (Figure S9) ----
write.csv(daylight_results, file.path(csv_dir, "amplitude_by_daylight.csv"), row.names = FALSE)

cat("All results CSVs saved to:", csv_dir, "\n")
```

## Sensitivity: December–January average as winter reference (individual-level)

```{r winter-avg-sensitivity}
# Holiday-related behavior is concentrated in late December. To confirm the
# winter peak is not December-specific, we recompute per-person seasonal
# amplitude using the mean of December and January as the winter reference
# (instead of December alone) and re-test age and sex group differences.
# This is the individual-level (no-spline) analog of Fig 1B/D and Fig 2B/D,
# computed on the same person-mean-centered scale (RHR_within / HRV_within),
# so values are directly comparable to the December-only estimates.

# ---- Winter = mean(December, January); Summer = August ----
df_winteravg <- df_month %>%
  filter(MONTH_NUM %in% c(1, 8, 12)) %>%
  mutate(SEASON = if_else(MONTH_NUM == 8, "August", "Winter")) %>%
  group_by(USER_ID, SEASON) %>%
  summarise(
    mean_rhr        = mean(RHR_within, na.rm = TRUE),
    mean_hrv        = mean(HRV_within, na.rm = TRUE),
    AGE_BUCKET      = first(AGE_BUCKET),
    AGE_2023        = first(AGE_2023),
    BMI             = first(BMI),
    GENDER          = first(GENDER),
    RHR_person_mean = first(RHR_person_mean),
    HRV_person_mean = first(HRV_person_mean),
    .groups = "drop"
  )

# RHR amplitude = Winter - August (winter minus summer)
df_amp_rhr_wa <- df_winteravg %>%
  select(USER_ID, SEASON, mean_rhr, AGE_BUCKET, AGE_2023, BMI, GENDER, RHR_person_mean) %>%
  pivot_wider(names_from = SEASON, values_from = mean_rhr) %>%
  mutate(amplitude = Winter - August)

# HRV amplitude = August - Winter (summer minus winter)
df_amp_hrv_wa <- df_winteravg %>%
  select(USER_ID, SEASON, mean_hrv, AGE_BUCKET, AGE_2023, BMI, GENDER, HRV_person_mean) %>%
  pivot_wider(names_from = SEASON, values_from = mean_hrv) %>%
  mutate(amplitude = August - Winter)

# ---- Per-age summaries (compare against the December-only values) ----
summarise_amp_wa <- function(d) {
  d %>%
    group_by(AGE_BUCKET) %>%
    summarise(
      mean_amplitude = mean(amplitude, na.rm = TRUE),
      sd_amplitude   = sd(amplitude, na.rm = TRUE),
      n              = n(),
      mean_sd        = paste0(round(mean_amplitude, 2), " ± ", round(sd_amplitude, 2)),
      .groups = "drop"
    )
}
amp_age_rhr_wa <- summarise_amp_wa(df_amp_rhr_wa) %>% mutate(Outcome = "RHR", Winter_ref = "Dec-Jan avg")
amp_age_hrv_wa <- summarise_amp_wa(df_amp_hrv_wa) %>% mutate(Outcome = "HRV", Winter_ref = "Dec-Jan avg")
cat("\n--- Individual-level amplitude by age (Dec-Jan average winter reference) ---\n")
print(amp_age_rhr_wa); print(amp_age_hrv_wa)

# ---- Age group differences (mirrors Fig 1B/D adjustment set) ----
model_amp_rhr_wa <- lm(amplitude ~ AGE_BUCKET + GENDER + BMI + RHR_person_mean, data = df_amp_rhr_wa)
model_amp_hrv_wa <- lm(amplitude ~ AGE_BUCKET + GENDER + BMI + HRV_person_mean, data = df_amp_hrv_wa)
tukey_age_rhr_wa <- as.data.frame(pairs(emmeans(model_amp_rhr_wa, ~ AGE_BUCKET), adjust = "tukey")) %>% mutate(Outcome = "RHR")
tukey_age_hrv_wa <- as.data.frame(pairs(emmeans(model_amp_hrv_wa, ~ AGE_BUCKET), adjust = "tukey")) %>% mutate(Outcome = "HRV")
cat("\n--- Age pairwise contrasts (Dec-Jan average winter reference) ---\n")
print(tukey_age_rhr_wa); print(tukey_age_hrv_wa)

# ---- Sex group differences (mirrors Fig 2B/D adjustment set) ----
model_amp_rhr_sex_wa <- lm(amplitude ~ GENDER + splines::ns(AGE_2023, df = 3) + BMI + RHR_person_mean, data = df_amp_rhr_wa)
model_amp_hrv_sex_wa <- lm(amplitude ~ GENDER + splines::ns(AGE_2023, df = 3) + BMI + HRV_person_mean, data = df_amp_hrv_wa)
tukey_sex_rhr_wa <- as.data.frame(pairs(emmeans(model_amp_rhr_sex_wa, ~ GENDER), adjust = "tukey")) %>% mutate(Outcome = "RHR")
tukey_sex_hrv_wa <- as.data.frame(pairs(emmeans(model_amp_hrv_sex_wa, ~ GENDER), adjust = "tukey")) %>% mutate(Outcome = "HRV")
cat("\n--- Sex pairwise contrasts (Dec-Jan average winter reference) ---\n")
print(tukey_sex_rhr_wa); print(tukey_sex_hrv_wa)

# ---- Save ----
write.csv(bind_rows(amp_age_rhr_wa, amp_age_hrv_wa),
          file.path(csv_dir, "individual_amplitude_winteravg_by_age.csv"), row.names = FALSE)
write.csv(bind_rows(tukey_age_rhr_wa, tukey_age_hrv_wa),
          file.path(csv_dir, "individual_amplitude_winteravg_tukey_age.csv"), row.names = FALSE)
write.csv(bind_rows(tukey_sex_rhr_wa, tukey_sex_hrv_wa),
          file.path(csv_dir, "individual_amplitude_winteravg_tukey_sex.csv"), row.names = FALSE)
cat("\nWinter-average sensitivity CSVs saved to:", csv_dir, "\n")
```